### 2022-02-14 README.Rmd
README working

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# melt

<!-- badges: start -->
[![R-CMD-check](https://github.com/markean/melt/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/markean/melt/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/melt)](https://CRAN.R-project.org/package=melt)
<!-- badges: end -->

The goal of melt is to ...

## Installation

You can install the development version of melt from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("markean/melt")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(melt)
## basic example code
```

What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:

```{r cars}
summary(cars)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.

### 2022-02-14 README.Rmd
README added & testthat added (empty curretly)

@@
 
-# melt
-
+# melt - Multiple Empirical Likelihood Tests
 <!-- badges: start -->
+[![CRAN status](https://www.r-pkg.org/badges/version/melt)](https://CRAN.R-project.org/package=melt)
 [![R-CMD-check](https://github.com/markean/melt/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/markean/melt/actions/workflows/R-CMD-check.yaml)
-[![CRAN status](https://www.r-pkg.org/badges/version/melt)](https://CRAN.R-project.org/package=melt)
 <!-- badges: end -->
 
-The goal of melt is to ...
+## Overview
+The R package **melt** provides a unified framework for data analysis with empirical likelihood methods.
+A collection of functions are available for basic regression analysis and hypothesis testing.
+Much of its functionality and syntax are designed to mimic the corresponding base R functions.
+The core routines are written in C++ and utilize OpenMP for parallelization.
 
 ## Installation
-
-You can install the development version of melt from [GitHub](https://github.com/) with:
-
-``` r
+You can install the latest stable release from [CRAN](https://cran.r-project.org/package=melt).
+```{r eval = FALSE}
+install.packages("melt", dependencies = TRUE)
+```
+You can install the latest development version from
+[Github](https://github.com/markean/melt).
+```{r eval = FALSE}
 # install.packages("devtools")
@@
 
-## Example
+## Usage
+```{r eval = FALSE}
+library(melt)
+# one sample test (mean)
+el_mean(par = 0, x = rnorm(n = 100))
 
-This is a basic example which shows you how to solve a common problem:
+# linear regression
+fit <- el_lm(formula = mpg ~ wt, data = mtcars)
+summary(fit)
 
-```{r example}
-library(melt)
-## basic example code
+# analysis of variance
+el_aov(formula = Sepal.Length ~ Species, data = iris)
 ```
-
-What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:
-
-```{r cars}
-summary(cars)
-```
-
-You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.
-`devtools::build_readme()` is handy for this.
-You could also use GitHub Actions to re-render `README.Rmd` every time you push.
-An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>.
-
-You can also embed plots, for example:
-
-```{r pressure, echo = FALSE}
-plot(pressure)
-```
-
-In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.

### 2022-02-14 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(melt)
+
+test_check("melt")

### 2022-02-22 src/EL.cpp
README now presents the outputs. EL2 class uses gradient for convergence check. This is experimental. Now, the version is 1.2.0.9000

@@ -111,10 +111,16 @@ EL2::EL2(const Eigen::Ref<const Eigen::VectorXd>& par0,
 }
 
 /* Constructor for weighted EL2 class (evaluation)
- * Last updated: 02/16/21
+ * Last updated: 02/21/21
  *
  * abstol for gamma should be reconsidered.
  * Perhaps, use another optim parameter such as step size tolerance.
+ *
+ * l2 norm of the gradient of nlogLR is
+ * (g.array().colwise() * log_tmp.dplog).colwise().sum().matrix().norm(), where
+ * the log_tmp is computed with the current value of lambda. Consider moving
+ * toward the real gradient descent than evaluating nlogLR directly. It could be
+ * slower.
  */
 EL2::EL2(const Eigen::Ref<const Eigen::VectorXd>& par0,
          const Eigen::Ref<const Eigen::MatrixXd>& x,
@@ -138,6 +144,13 @@ EL2::EL2(const Eigen::Ref<const Eigen::VectorXd>& par0,
   while (!convergence && iterations != maxit && nlogLR <= threshold) {
     // plog class
     PSEUDO_LOG log_tmp(Eigen::VectorXd::Ones(g.rows()) + g * lambda, w);
+    // convergence check
+    if ((g.array().colwise() * log_tmp.dplog).colwise().sum().matrix().norm() <
+      abstol) {
+      nlogLR = log_tmp.plog_sum;
+      convergence = true;
+      break;
+    }
     // J matrix
     const Eigen::MatrixXd J = g.array().colwise() * log_tmp.sqrt_neg_d2plog;
     // propose new lambda by NR method with least square
@@ -162,15 +175,7 @@ EL2::EL2(const Eigen::Ref<const Eigen::VectorXd>& par0,
      * hull constraint), terminate the maximization with the current values
      * without further updates.
      */
-    if (gamma < abstol) {
-      nlogLR = log_tmp.plog_sum;
-      break;
-    }
-    // Otherwise, update lambda and check for convergence
     lambda += gamma * step;
-    if (nlogLR - log_tmp.plog_sum < abstol) {
-      convergence = true;
-    }
     ++iterations;
   }
 }

### 2022-03-10 src/utils_gbd.cpp
gbd related functions kind of sandboxed

@@ -1,5 +1,101 @@
 #include "utils_gbd.h"
 
+EL_deprecated::EL_deprecated(const Eigen::Ref<const Eigen::MatrixXd>& g,
+                             const double threshold,
+                             const int maxit,
+                             const double abstol) {
+  // maximization
+  lambda = (g.transpose() * g).ldlt().solve(g.colwise().sum());
+  iterations = 0;
+  convergence = false;
+  while (!convergence && iterations != maxit) {
+    // plog class
+    PSEUDO_LOG_deprecated log_tmp(Eigen::VectorXd::Ones(g.rows()) + g * lambda);
+    // J matrix
+    const Eigen::MatrixXd J = g.array().colwise() * log_tmp.sqrt_neg_d2plog;
+    // prpose new lambda by NR method with least square
+    Eigen::VectorXd step =
+      (J.transpose() * J).ldlt().solve(
+          J.transpose() * (log_tmp.dplog / log_tmp.sqrt_neg_d2plog).matrix());
+    // update function value
+    nlogLR =
+      PSEUDO_LOG_deprecated::sum(Eigen::VectorXd::Ones(g.rows()) + g * (lambda + step));
+    // step halving to ensure validity
+    while (nlogLR < log_tmp.plog_sum) {
+      step /= 2;
+      nlogLR =
+        PSEUDO_LOG_deprecated::sum(Eigen::VectorXd::Ones(g.rows()) + g * (lambda + step));
+    }
+    // update lambda
+    lambda += step;
+
+    // check convex hull constraint(stop if larger than threshold)
+    if (nlogLR > threshold) {
+      break;
+    }
+
+    // convergence check
+    if (nlogLR - log_tmp.plog_sum < abstol) {
+      convergence = true;
+    } else {
+      ++iterations;
+    }
+  }
+}
+
+PSEUDO_LOG_deprecated::PSEUDO_LOG_deprecated(Eigen::VectorXd&& x) {
+  static const double n = static_cast<double>(x.size());
+  static const double a0 = 1.0 / n;
+  static const double a1 = -std::log(n) - 1.5;
+  static const double a2 = 2.0 * n;
+  static const double a3 = -0.5 * n * n;
+
+  dplog.resize(x.size());
+  sqrt_neg_d2plog.resize(x.size());
+  plog_sum = 0;
+
[…]

### 2022-03-10 src/utils_gbd.h

@@ -1,8 +1,39 @@
 #ifndef EL_UTILS_gbd_H_
 #define EL_UTILS_gbd_H_
 
-#include "EL.h"
 #include "utils.h"
+
+struct minEL {
+  Eigen::VectorXd par;
+  Eigen::VectorXd lambda;
+  double nlogLR;
+  int iterations;
+  bool convergence;
+};
+
+class EL_deprecated {
+public:
+  Eigen::VectorXd lambda;
+  double nlogLR;
+  int iterations;
+  bool convergence;
+
+  EL_deprecated(const Eigen::Ref<const Eigen::MatrixXd>& g,
+                const double threshold,
+                const int maxit = 100,
+                const double abstol = 1e-8);
+};
+
+class PSEUDO_LOG_deprecated {
+public:
+  Eigen::ArrayXd dplog;
+  Eigen::ArrayXd sqrt_neg_d2plog;
+  double plog_sum;
+
+  PSEUDO_LOG_deprecated(Eigen::VectorXd&& x);
+  static double sum(Eigen::VectorXd&& x);
+  static Eigen::ArrayXd dp(Eigen::VectorXd&& x);
+};
 
 Eigen::MatrixXd g_gbd(const Eigen::Ref<const Eigen::VectorXd>& theta,
                       const Eigen::Ref<const Eigen::MatrixXd>& x,
@@ -18,18 +49,18 @@ Eigen::VectorXd lambda2theta_gbd(const Eigen::Ref<const Eigen::VectorXd>& lambda
                                  const double gamma);
 
 void lambda2theta_void(
-        const Eigen::Ref<const Eigen::VectorXd>& lambda,
-        Eigen::Ref<Eigen::VectorXd> theta,
-        const Eigen::Ref<const Eigen::MatrixXd>& g,
-        const Eigen::Ref<const Eigen::MatrixXd>& c,
-        const double gamma);
+                const Eigen::Ref<const Eigen::VectorXd>& lambda,
+                Eigen::Ref<Eigen::VectorXd> theta,
+                const Eigen::Ref<const Eigen::MatrixXd>& g,
+                const Eigen::Ref<const Eigen::MatrixXd>& c,
+                const double gamma);
 
 Eigen::VectorXd approx_lambda_gbd(
-        const Eigen::Ref<const Eigen::MatrixXd>& g0,
-        const Eigen::Ref<const Eigen::MatrixXd>& c,
[…]

### 2022-03-10 src/utils_pairwise.h

@@ -13,35 +13,35 @@
 
 
 std::vector<std::array<int, 2>> comparison_pairs(
-    const int p, const int control);
+        const int p, const int control);
 
 std::array<double, 2> pair_confidence_interval_gbd(
-    const Eigen::Ref<const Eigen::VectorXd>& theta0,
-    const Eigen::Ref<const Eigen::MatrixXd>& x,
-    const Eigen::Ref<const Eigen::MatrixXd>& c,
-    const Eigen::Ref<const Eigen::MatrixXd>& lhs,
-    const double threshold,
-    const double init,
-    const double cutoff);
+        const Eigen::Ref<const Eigen::VectorXd>& theta0,
+        const Eigen::Ref<const Eigen::MatrixXd>& x,
+        const Eigen::Ref<const Eigen::MatrixXd>& c,
+        const Eigen::Ref<const Eigen::MatrixXd>& lhs,
+        const double threshold,
+        const double init,
+        const double cutoff);
 
 Eigen::ArrayXd bootstrap_statistics_pairwise_AMC(
-    const Eigen::Ref<const Eigen::MatrixXd>& x,
-    const Eigen::Ref<const Eigen::MatrixXd>& c,
-    const int k,
-    const std::vector<std::array<int, 2>>& pairs,
-    const int B,
-    const double level);
+        const Eigen::Ref<const Eigen::MatrixXd>& x,
+        const Eigen::Ref<const Eigen::MatrixXd>& c,
+        const int k,
+        const std::vector<std::array<int, 2>>& pairs,
+        const int B,
+        const double level);
 
 Eigen::ArrayXd bootstrap_statistics_pairwise_NB(
-    const Eigen::Ref<const Eigen::MatrixXd>& x,
-    const Eigen::Ref<const Eigen::MatrixXd>& c,
-    const int k,
-    const std::vector<std::array<int, 2>>& pairs,
-    const int B,
-    const double level,
-    const int nthread,
-    const bool progress,
-    const double threshold,
-    const int maxit,
-    const double abstol);
+        const Eigen::Ref<const Eigen::MatrixXd>& x,
+        const Eigen::Ref<const Eigen::MatrixXd>& c,
+        const int k,
+        const std::vector<std::array<int, 2>>& pairs,
+        const int B,
+        const double level,
+        const int nthread,
+        const bool progress,
+        const double threshold,
+        const int maxit,
+        const double abstol);
 #endif

### 2022-03-20 src/EL.h
EL class redesigned. convergence criterion updated. threshold is available.

@@ -3,7 +3,7 @@
 
 #include "eigen_config.h"
 #include <RcppEigen.h>
-// #include <cmath>
+#include <cmath>
 #include "utils.h"
 
 class EL
@@ -11,25 +11,21 @@ class EL
 private:
   const int n;
 public:
-  Eigen::VectorXd lambda;
-  double nlogLR = 0;
-  int iterations = 1;
-  bool convergence = false;
+  Eigen::VectorXd l;
+  double nlogLR;
+  int iterations{1};
+  bool convergence{false};
 
   EL(const Eigen::Ref<const Eigen::MatrixXd>& g,
      const int maxit,
-     const double abstol,
+     const double tol,
      const double threshold);
 
   EL(const Eigen::Ref<const Eigen::MatrixXd>& g,
-     const Eigen::Ref<const Eigen::ArrayXd>& w,
+     const Eigen::Ref<const Eigen::VectorXd>& w,
      const int maxit,
-     const double abstol,
+     const double tol,
      const double threshold);
-
-  // // log probability
-  // Eigen::ArrayXd log_prob(const Eigen::Ref<const Eigen::MatrixXd>& g,
-  //                         const Eigen::Ref<const Eigen::ArrayXd>& w) const;
 };
 
 class EL2
@@ -56,7 +52,7 @@ public:
   EL2(const std::string method,
       const Eigen::Ref<const Eigen::VectorXd>& par0,
       const Eigen::Ref<const Eigen::MatrixXd>& x,
-      const Eigen::Ref<const Eigen::ArrayXd>& w,
+      const Eigen::Ref<const Eigen::VectorXd>& w,
       const int maxit,
       const double abstol,
       const double threshold);
@@ -87,13 +83,12 @@ public:
   Eigen::ArrayXd sqrt_neg_d2plog;
   double plog_sum = 0;
 
-  // PSEUDO_LOG(const Eigen::Ref<const Eigen::VectorXd>& x);
   PSEUDO_LOG(Eigen::VectorXd&& x);
-  PSEUDO_LOG(Eigen::VectorXd&& x, Eigen::ArrayXd&& w);
-  static Eigen::ArrayXd plog(Eigen::ArrayXd&& x);
+  PSEUDO_LOG(Eigen::VectorXd&& x, const Eigen::Ref<const Eigen::VectorXd>& w);
+  static Eigen::ArrayXd plog(Eigen::VectorXd&& x);
[…]

### 2022-03-22 README.Rmd
README now presents the outputs. EL2 class uses gradient for convergence check. This is experimental. Now, the version is 1.2.0.9000
gbd related functions kind of sandboxed
EL class redesigned. convergence criterion updated. threshold is available.
abstol replaced by tol

@@
 ```{r eval = FALSE}
-install.packages("melt", dependencies = TRUE)
+install.packages("melt")
 ```
@@
 ## Usage
-```{r eval = FALSE}
+```{r eval = TRUE}
 library(melt)

### 2022-03-22 R/RcppExports.R

@@ -1,43 +1,43 @@
 # Generated by using Rcpp::compileAttributes() -> do not edit by hand
 # Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
 
-EL_eval <- function(g, maxit, abstol, threshold) {
-    .Call(`_melt_EL_eval`, g, maxit, abstol, threshold)
+EL_eval <- function(g, maxit, abstol, th) {
+    .Call(`_melt_EL_eval`, g, maxit, abstol, th)
 }
 
-WEL_eval <- function(g, w, maxit, abstol, threshold) {
-    .Call(`_melt_WEL_eval`, g, w, maxit, abstol, threshold)
+WEL_eval <- function(g, w, maxit, abstol, th) {
+    .Call(`_melt_WEL_eval`, g, w, maxit, abstol, th)
 }
 
 ELtest <- function(x, c, lhs, rhs, threshold, maxit = 1e4L, abstol = 1e-8) {
     .Call(`_melt_ELtest`, x, c, lhs, rhs, threshold, maxit, abstol)
 }
 
-EL_lm <- function(data, intercept, maxit, abstol, threshold) {
-    .Call(`_melt_EL_lm`, data, intercept, maxit, abstol, threshold)
+EL_lm <- function(data, intercept, maxit, abstol, th) {
+    .Call(`_melt_EL_lm`, data, intercept, maxit, abstol, th)
 }
 
-EL_mean <- function(par, x, maxit, tol, threshold) {
-    .Call(`_melt_EL_mean`, par, x, maxit, tol, threshold)
+EL_mean <- function(par, x, maxit, tol, th) {
+    .Call(`_melt_EL_mean`, par, x, maxit, tol, th)
 }
 
-WEL_mean <- function(par, x, w, maxit, tol, threshold) {
-    .Call(`_melt_WEL_mean`, par, x, w, maxit, tol, threshold)
+WEL_mean <- function(par, x, w, maxit, tol, th) {
+    .Call(`_melt_WEL_mean`, par, x, w, maxit, tol, th)
 }
 
-EL_confint <- function(x, type, init, cutoff, idx, maxit, abstol, threshold) {
-    .Call(`_melt_EL_confint`, x, type, init, cutoff, idx, maxit, abstol, threshold)
+EL_confint <- function(x, type, init, cutoff, idx, maxit, abstol, th) {
+    .Call(`_melt_EL_confint`, x, type, init, cutoff, idx, maxit, abstol, th)
 }
 
[…]

### 2022-03-22 R/el_mean.R

@@ -49,7 +49,7 @@ el_mean <- function(par, x, weights = NULL, control = list())
   # check control
   optcfg <- check_control(control)
   if (is.null(weights)) {
-    out <- EL_mean(par, mm, optcfg$maxit, optcfg$abstol, optcfg$threshold)
+    out <- EL_mean(par, mm, optcfg$maxit, optcfg$tol, optcfg$threshold)
   } else {
     if (!is.numeric(weights))
       stop("'weights' must be a numeric vector")
@@ -61,7 +61,7 @@ el_mean <- function(par, x, weights = NULL, control = list())
     if (length(w) != NROW(mm))
       stop("'g' and 'weights' have incompatible dimensions")
     w <- (NROW(mm) / sum(w)) * w
-    out <- WEL_mean(par, mm, w, optcfg$maxit, optcfg$abstol, optcfg$threshold)
+    out <- WEL_mean(par, mm, w, optcfg$maxit, optcfg$tol, optcfg$threshold)
   }
   out$data.matrix <- mm
   out

### 2022-03-22 src/EL.h
EL2 classed replaced by EL class. function pointer is used for setting methods such as estimating function and gradient function.

@@ -6,84 +6,103 @@
 #include <cmath>
 #include "utils.h"
 
-class EL2
+class EL
 {
-private:
-  const std::string type;
-  Eigen::VectorXd par;
-  const int maxit;
-  const double tol;
-  const double threshold;
-  const int n;
 public:
-  Eigen::VectorXd l;
-  double nlogLR{0};
-  int iterations{1};
-  bool convergence{false};
+  // members
+  Eigen::VectorXd l;  // Lagrange multiplier
+  double nllr{0};     // negative log likelihood ratio
+  int iter{0};        // iterations performed in optimization
+  bool conv{false};   // convergence status
 
-  void setup(const Eigen::Ref<const Eigen::MatrixXd>& g);
-
-  void setup(const Eigen::Ref<const Eigen::MatrixXd>& g,
-             const Eigen::Ref<const Eigen::VectorXd>& w);
-
+  // constructors
   // direct evaluation
-  EL2(const Eigen::Ref<const Eigen::MatrixXd>& g,
-      const int maxit,
-      const double tol,
-      const double threshold);
-
+  EL(const Eigen::Ref<const Eigen::MatrixXd>& g,
+     const int maxit,
+     const double tol,
+     const double th);
   // direct evaluation (weighted)
-  EL2(const Eigen::Ref<const Eigen::MatrixXd>& g,
-      const Eigen::Ref<const Eigen::VectorXd>& w,
-      const int maxit,
-      const double tol,
-      const double threshold);
-
+  EL(const Eigen::Ref<const Eigen::MatrixXd>& g,
+     const Eigen::Ref<const Eigen::VectorXd>& w,
+     const int maxit,
+     const double tol,
+     const double th);
   // evaluation
-  EL2(const std::string method,
-      const Eigen::Ref<const Eigen::VectorXd>& par0,
-      const Eigen::Ref<const Eigen::MatrixXd>& x,
-      const int maxit,
-      const double tol,
-      const double threshold);
-
+  EL(const std::string method,
+     const Eigen::Ref<const Eigen::VectorXd>& par0,
+     const Eigen::Ref<const Eigen::MatrixXd>& x,
+     const int maxit,
+     const double tol,
+     const double th);
   // evaluation (weighted)
-  EL2(const std::string method,
-      const Eigen::Ref<const Eigen::VectorXd>& par0,
[…]

### 2022-03-22 src/utils.cpp

@@ -1,32 +1,34 @@
 #include "utils.h"
 
-double th_nlogLR(const int p, const Rcpp::Nullable<double> threshold) {
-  return (threshold.isNull())? 20.0 * p : Rcpp::as<double>(threshold);
+double th_nloglr(const int p, const Rcpp::Nullable<double> th)
+{
+  return (th.isNull())? 20.0 * p : Rcpp::as<double>(th);
 }
 
-Eigen::MatrixXd g_mean(const Eigen::Ref<const Eigen::VectorXd>& par,
-                       const Eigen::Ref<const Eigen::MatrixXd>& x) {
+Eigen::MatrixXd g_mean(const Eigen::Ref<const Eigen::MatrixXd>& x,
+                       const Eigen::Ref<const Eigen::VectorXd>& par)
+{
   return x.rowwise() - par.transpose();
 }
 
-Eigen::MatrixXd g_lm(const Eigen::Ref<const Eigen::VectorXd>& par,
-                     const Eigen::Ref<const Eigen::MatrixXd>& data) {
+Eigen::MatrixXd g_lm(const Eigen::Ref<const Eigen::MatrixXd>& x,
+                     const Eigen::Ref<const Eigen::VectorXd>& par)
+{
   // const Eigen::VectorXd y = data.col(0);
   // const Eigen::MatrixXd x = data.rightCols(data.cols() - 1);
   // return x.array().colwise() * (y - x * beta).array();
-  return data.rightCols(data.cols() - 1).array().colwise() *
-    (data.col(0) - data.rightCols(data.cols() - 1) * par).array();
+  return x.rightCols(x.cols() - 1).array().colwise() *
+    (x.col(0) - x.rightCols(x.cols() - 1) * par).array();
 }
 
-Eigen::VectorXd gr_nlogLR_lm(
+Eigen::VectorXd gr_nloglr_lm(
     const Eigen::Ref<const Eigen::VectorXd>& lambda,
     const Eigen::Ref<const Eigen::MatrixXd>& g,
-    const Eigen::Ref<const Eigen::MatrixXd>& data) {
+    const Eigen::Ref<const Eigen::MatrixXd>& data)
+{
   const Eigen::MatrixXd x = data.rightCols(data.cols() - 1);
   const Eigen::ArrayXd denominator =
     Eigen::VectorXd::Ones(g.rows()) + g * lambda;
-  // const Eigen::MatrixXd gradient =
[…]

### 2022-03-30 README.Rmd
EL2 classed replaced by EL class. function pointer is used for setting methods such as estimating function and gradient function.
README, DESCRIPTION updated

@@
 library(melt)
-# one sample test (mean) 
-el_mean(par = 0, x = rnorm(n = 100))
+# one sample test for mean
+fit1 <- el_mean(par = 0, x = rnorm(n = 100))
+confint(fit1)
 
 # linear regression
-fit <- el_lm(formula = mpg ~ wt, data = mtcars)
-summary(fit)
+fit2 <- el_lm(formula = mpg ~ wt, data = mtcars)
+summary(fit2)
+confint(fit2)
 
 # analysis of variance 
-el_aov(formula = Sepal.Length ~ Species, data = iris)
+data("clothianidin")
+fit3 <- el_lm(clo ~ -1 + trt, clothianidin)
+summary(fit3)
+confint(fit3)
+
+# test of no treatment effect
+lhs <- matrix(c(1, -1, 0, 0,
+                0, 1, -1, 0,
+                0, 0, 1, -1), byrow = TRUE, nrow = 3)
+lht(fit3, lhs)
 ```

### 2022-03-30 README.Rmd
codecov added for CI. default th set to 200*p. generic logLik added.

@@
 [![R-CMD-check](https://github.com/markean/melt/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/markean/melt/actions/workflows/R-CMD-check.yaml)
+[![Codecov test coverage](https://codecov.io/gh/markean/melt/branch/master/graph/badge.svg)](https://app.codecov.io/gh/markean/melt?branch=master)
 <!-- badges: end -->

### 2022-03-30 R/checker.R

@@ -1,3 +1,4 @@
+#' @noRd
 check_control <- function(control = list()) {
   ctrl <- list(maxit = 100L, tol = 1e-06, th = NULL)
   nctrl <- names(ctrl)

### 2022-03-30 R/el_mean.R

@@ -11,7 +11,7 @@
 #' @param control A list of control parameters. See ‘Details’ in
 #'   \code{\link{el_eval}}.
 #' @return A list with class \code{"el_test"} as described in
-#'   \code{\link{el_eval}}.
+#'   \code{\link{lht}}.
 #' @references Glenn, N.L., and Yichuan Zhao. 2007.
 #'   “Weighted Empirical Likelihood Estimates and Their Robustness Properties.”
 #'   Computational Statistics & Data Analysis 51 (10): 5130–41.
@@ -19,19 +19,19 @@
 #' @references Owen, Art. 1990. “Empirical Likelihood Ratio Confidence Regions.”
 #'   The Annals of Statistics 18 (1).
 #'   \doi{10.1214/aos/1176347494}.
-#' @seealso \link{el_eval}
+#' @seealso \link{el_eval}, \link{lht}
 #' @examples
-#' ## scalar mean
+#' # scalar mean
 #' par <- 0
 #' x <- rnorm(100)
 #' el_mean(par, x)
 #'
-#' ## vector mean
+#' # vector mean
 #' par <- c(0, 0)
 #' x <- matrix(rnorm(100L), ncol = 2)
 #' el_mean(par, x)
 #'
-#' ## weighted EL
+#' # weighted EL
 #' par <- c(0, 0)
 #' x <- matrix(rnorm(100), ncol = 2)
 #' w <- rep(c(1, 2), each = 25)

### 2022-03-30 src/EL.cpp

@@ -164,6 +164,8 @@ Eigen::ArrayXd EL::log_wprob(const Eigen::Ref<const Eigen::MatrixXd>& x,
 }
 
 
+
+
 Eigen::ArrayXd EL::logp(const Eigen::Ref<const Eigen::MatrixXd>& x) const {
   const Eigen::MatrixXd g = g_fcn(x, par);
   return  -log(n) - PSEUDO_LOG::plog(Eigen::VectorXd::Ones(n) + g * l);
@@ -182,95 +184,24 @@ Eigen::ArrayXd EL::logp_g(const Eigen::Ref<const Eigen::MatrixXd>& g) const {
 
 Eigen::ArrayXd EL::logp_g(const Eigen::Ref<const Eigen::MatrixXd>& g,
                           const Eigen::Ref<const Eigen::ArrayXd>& w) const {
-  // return w.log() -log(n) - PSEUDO_LOG::plog(Eigen::VectorXd::Ones(n) + g * l);
   return PSEUDO_LOG::plog(w) - log(n) -
     PSEUDO_LOG::plog(Eigen::VectorXd::Ones(n) + g * l);
 }
 
-// /* Constructor for EL class (minimization)
-//  * Last updated: 03/21/21
-//  */
-// EL::EL(const std::string method,
-//        const Eigen::Ref<const Eigen::VectorXd>& par0,
-//        const Eigen::Ref<const Eigen::MatrixXd>& x,
-//        const Eigen::Ref<const Eigen::MatrixXd>& lhs,
-//        const Eigen::Ref<const Eigen::VectorXd>& rhs,
-//        const int maxit,
-//        const double tol,
-//        const double th)
-//   : maxit{maxit},
-//     tol{tol},
-//     th{th},
-//     n{static_cast<int>(x.rows())},
-//     g_fcn{set_g_fcn(method)},
-//     gr_fcn{set_gr_fcn(method)}
-// {
-//   /// initialization ///
-//   // orthogonal projection matrix
-//   const Eigen::MatrixXd proj =
-//     Eigen::MatrixXd::Identity(lhs.cols(), lhs.cols()) -
-//     lhs.transpose() * (lhs * lhs.transpose()).inverse() * lhs;
-//   // parameter (constraint imposed)
-//   par = proj * par0 + lhs.transpose() * (lhs * lhs.transpose()).inverse() * rhs;
-//   // estimating function
[…]

### 2022-03-30 src/lm.cpp

@@ -23,15 +23,12 @@ Rcpp::List lm_(const Eigen::Map<Eigen::MatrixXd>& data,
   lhs.col(0) = Eigen::MatrixXd::Zero(p - 1, 1);
   lhs.rightCols(p - 1) = Eigen::MatrixXd::Identity(p - 1, p - 1);
   const Eigen::VectorXd rhs = Eigen::VectorXd::Zero(p - 1);
-  // const EL el =
-  //   (intercept && p > 1)?
-  //   EL("lm", bhat, data, lhs, rhs, maxit, tol, th_nloglr(p - 1, th)) :
-  //   EL("lm", Eigen::MatrixXd::Zero(1, p), data , maxit, tol, th_nloglr(p, th));
   const MINEL el =
     (intercept && p > 1)?
     MINEL("lm", bhat, data, lhs, rhs, maxit, tol, th_nloglr(p - 1, th)) :
     MINEL("lm", bhat, data, Eigen::MatrixXd::Identity(p, p),
           Eigen::VectorXd::Zero(p), maxit, tol,th_nloglr(p, th));
+  const int df = (intercept && p > 1)? p - 1 : p;
 
   // test each coefficient
   Rcpp::NumericVector chisq_val(p);
@@ -63,7 +60,10 @@ Rcpp::List lm_(const Eigen::Map<Eigen::MatrixXd>& data,
         Rcpp::Named("convergence") = conv)),
     Rcpp::Named("npar") = p,
     Rcpp::Named("log.prob") = el.logp(data),
+    Rcpp::Named("loglik") = el.loglik(),
     Rcpp::Named("coefficients") = bhat,
+    // Rcpp::Named("statistic") = ,
+    Rcpp::Named("df") = df,
     Rcpp::Named("residuals") = resid,
     Rcpp::Named("fitted.values") = fit_val);
   result.attr("class") = Rcpp::CharacterVector({"el_lm", "el_test"});
@@ -101,6 +101,7 @@ Rcpp::List lm_w_(const Eigen::Map<Eigen::MatrixXd>& data,
     MINEL("lm", bhat, data, w, lhs, rhs, maxit, tol, th_nloglr(p - 1, th)) :
     MINEL("lm", bhat, data, w, Eigen::MatrixXd::Identity(p, p),
           Eigen::VectorXd::Zero(p), maxit, tol, th_nloglr(p, th));
+  const int df = (intercept && p > 1)? p - 1 : p;
 
   // test each coefficient
[…]

### 2022-03-30 tests/testthat/test-el_eval.R

@@ -6,7 +6,7 @@ test_that("convergence check", {
     el_eval(x - par, control =
               list(maxit = 20L, tol = 1e-08, th = 1e+10))$optim$convergence
   }
-  expect_true(all(sapply(grid, conv)))
+  expect_true(all(vapply(grid, conv, FUN.VALUE = logical(1))))
 })
 
 test_that("probabilities add up to 1", {
@@ -18,15 +18,37 @@ test_that("probabilities add up to 1", {
   expect_equal(sum(exp(fit$log.prob)), 1)
 })
 
-# test_that("probabilities add up to 1 (weighted)", {
-#   skip_on_os("windows", arch = "i386")
-#   x <- rnorm(10)
-#   par <- (max(x) + min(x)) / 2
-#   w <- 1 + runif(10, min = -0.2, max = 0.2)
-#   optcfg <- list(maxit = 200L, tol = 1e-08, th = 1e+10)
-#   fit <- el_eval(x - par, w, optcfg)
-#   expect_equal(sum(exp(fit$log.prob)), 1)
-# })
+test_that("probabilities add up to 1 (weighted)", {
+  skip_on_os("windows", arch = "i386")
+  x <- rnorm(10)
+  par <- (max(x) + min(x)) / 2
+  w <- 1 + runif(10, min = -0.5, max = 0.5)
+  optcfg <- list(maxit = 200L, tol = 1e-08, th = 1e+10)
+  fit <- el_eval(x - par, w, optcfg)
+  expect_equal(sum(exp(fit$log.prob)), 1)
+})
+
+test_that("loglik to loglr", {
+  skip_on_os("windows", arch = "i386")
+  n <- 10
+  x <- rnorm(n)
+  par <- (max(x) + min(x)) / 2
+  optcfg <- list(maxit = 200L, tol = 1e-08, th = 1e+10)
+  fit <- el_eval(x - par, control = optcfg)
+  expect_equal(fit$loglik + n * log(n), fit$optim$logLR)
+})
+
+test_that("loglik to loglr (weighted)", {
+  skip_on_os("windows", arch = "i386")
+  n <- 10
+  x <- rnorm(n)
+  par <- (max(x) + min(x)) / 2
+  w <- 1 + runif(n, min = -0.5, max = 0.5)
+  optcfg <- list(maxit = 200L, tol = 1e-08, th = 1e+10)
+  fit <- el_eval(x - par, w, optcfg)
+  w <- fit$weights
[…]

### 2022-03-30 R/el_aov.R
test coverage

@@ -95,6 +95,7 @@ el_aov <- function(formula, data, maxit = 1e04, abstol = 1e-8) {
   out
 }
 
+#' @noRd
 #' @export
 print.el_aov <- function(x, ...) {
   cat("\nCall:\n")

### 2022-03-30 README.Rmd
test coverage
vignette skeleton added.

@@
 A collection of functions are available for basic regression analysis and hypothesis testing.
-Much of its functionality and syntax are designed to mimic the corresponding base R functions.
-The core routines are written in C++ and utilize OpenMP for parallelization.
+Much of its functionality and syntax mimics the corresponding base R functions.
+The core computational routines are implemented with the 'Eigen' C++ library and 'RcppEigen' interface, with OpenMP for parallel computation.
+Additional functions are available for multiple testing for the analysis of experimental designs.
+Details of the testing procedures are given in [Kim et al.
+(2021)](https://arxiv.org/abs/2112.09206).
 

### 2022-03-30 R/el_aov.R

@@ -32,8 +32,7 @@ el_aov <- function(formula, data, maxit = 1e04, abstol = 1e-8) {
     # no other formula
     typeof(f$variables[[3]]) != "symbol" ||
       length(f$variables[[3]]) != 1
-  )
-  ) {
+  )) {
     stop("invalied model formula. specify formula as 'response ~ treatment'")
   }
 

### 2022-03-30 src/EL.cpp

@@ -143,29 +143,6 @@ EL::EL(const std::string method,
   set_el(g_fcn(x, par), w);
 }
 
-/* log probability for weighted EL class
- * Last updated: 03/09/21
- *
- */
-Eigen::ArrayXd EL::log_prob(const Eigen::Ref<const Eigen::MatrixXd>& x,
-                            const Eigen::Ref<const Eigen::ArrayXd>& w) const {
-  const Eigen::MatrixXd g = g_fcn(x, par);
-  return  w.log() - PSEUDO_LOG::plog(Eigen::VectorXd::Ones(n) + g * l);
-}
-
-/* log weighted probability for weighted EL class
- * Last updated: 03/09/21
- *
- */
-Eigen::ArrayXd EL::log_wprob(const Eigen::Ref<const Eigen::MatrixXd>& x,
-                             const Eigen::Ref<const Eigen::ArrayXd>& w) const {
-  const Eigen::MatrixXd g = g_fcn(x, par);
-  return  -w * PSEUDO_LOG::plog(Eigen::VectorXd::Ones(n) + g * l);
-}
-
-
-
-
 Eigen::ArrayXd EL::logp(const Eigen::Ref<const Eigen::MatrixXd>& x) const {
   const Eigen::MatrixXd g = g_fcn(x, par);
   return  -log(n) - PSEUDO_LOG::plog(Eigen::VectorXd::Ones(n) + g * l);

### 2022-03-30 tests/testthat/test-el_pairwise.R

@@ -1,0 +1,9 @@
+test_that("output", {
+  skip_on_os("windows", arch = "i386")
+  data("clothianidin")
+  out1 <- el_pairwise(clo ~ trt | blk, clothianidin, B = 500, progress = TRUE)
+  out2 <- el_pairwise(clo ~ trt | blk, clothianidin, control = "Naked",
+                      method = "NB", progress = TRUE, B = 2000, nthread = 2)
+  expect_output(print(out1))
+  expect_output(print(out2))
+})

### 2022-04-03 R/el_eval.R
1.4.0 release ready

@@ -88,11 +88,14 @@ el_eval <- function(g, weights, control = list()) {
 
   # check control
   optcfg <- check_control(control)
+  maxit <- optcfg$maxit
+  tol <- optcfg$tol
+  th <- optcfg$th
   if (missing(weights)) {
-    out <- eval_g_(mm, optcfg$maxit, optcfg$tol, optcfg$th)
+    out <- eval_g_(mm, maxit, tol, th)
   } else {
-    w <- check_weights(weights, NROW(mm))
-    out <- eval_g_w_(mm, w, optcfg$maxit, optcfg$tol, optcfg$th)
+    w <- check_weights(weights, nrow(mm))
+    out <- eval_g_w_(mm, w, maxit, tol, th)
     out$weights <- w
   }
   out

### 2022-04-03 R/lht.R

@@ -98,7 +98,6 @@ lht <- function(object, rhs = NULL, lhs = NULL, control = list()) {
   tol <- optcfg$tol
   th <- optcfg$th
   w <- object$weights
-
   if (is.null(lhs)) {
     if (is.null(w)) {
       out <- eval_(method, h$r, object$data.matrix, maxit, tol, th)
@@ -114,41 +113,6 @@ lht <- function(object, rhs = NULL, lhs = NULL, control = list()) {
                     h$r, maxit, tol, th)
     }
   }
-  # if (missing(lhs)) {
-  #   rhs <- check_rhs(rhs, p)
-  #   if (is.null(w)) {
-  #     out <- eval_(object$optim$method, rhs, object$data.matrix, maxit, tol, th)
-  #   } else {
-  #     out <- eval_w_(object$optim$method, rhs, object$data.matrix, w, maxit,
-  #                    tol, th)
-  #   }
-  # } else if (missing(rhs)) {
-  #   lhs <- as.matrix(lhs)
-  #   q <- check_hypothesis(lhs, p)
-  # }
-  #
-  #
-  # lhs <- as.matrix(lhs)
-  # q <- check_hypothesis(lhs, p)
-  #
-  # if (missing(rhs)) {
-  #   rhs <- rep(0, nrow(lhs))
-  # } else {
-  #   rhs <- as.vector(rhs)
-  #   if (!is.numeric(rhs) || !all(is.finite(rhs)))
-  #     stop("'rhs' must be a finite numeric vector")
-  #   if (length(rhs) != q)
-  #     stop("'lhs' and 'rhs' have incompatible dimensions")
-  # }
-  #
-  #
-  # if (is.null(object$weights)) {
-  #   out <- lht_(object$optim$method, object$coefficients, object$data.matrix,
-  #               lhs, rhs, optcfg$maxit, optcfg$tol, optcfg$th)
-  # } else {
-  #   out <- lht_w_(object$optim$method, object$coefficients, object$data.matrix,
-  #                 object$weights, lhs, rhs, optcfg$maxit, optcfg$tol, optcfg$th)
-  # }
   class(out) <- c("el_lht", "el_test")
   out
 }

### 2022-05-03 README.Rmd
CRAN 1.5.0 release ready.

@@
                 0, 0, 1, -1), byrow = TRUE, nrow = 3)
-lht(fit3, lhs)
+lht(fit3, lhs = lhs)
+
+# logistic regression
+fit4 <- el_glm(vs ~ mpg, data = mtcars, family = binomial)
+summary(fit4)
 ```

### 2022-05-03 R/AllGenerics.R

@@ -15,7 +15,7 @@
 #'   If \eqn{\textnormal{ELD}_i } is large, then the \eqn{i}th observation is an
 #'   influential point and can be inspected as a possible outlier. \code{eld}
 #'   computes \eqn{\textnormal{ELD}_i } for \eqn{i = 1, \dots, n }.
-#' @return An S4 object of class \linkS4class{ELD}.
+#' @return An object of class \linkS4class{ELD}.
 #' @references Lazar, Nicole A. 2005. “Assessing the Effect of Individual Data
 #'   Points on Inference From Empirical Likelihood.” Journal of Computational
 #'   and Graphical Statistics 14 (3): 626–42.
@@ -103,7 +103,7 @@ setGeneric("confint", function(object, parm, level = 0.95, ...)
 #'   \code{50}.
 #' @param control A list of control parameters set by \code{\link{el_control}}.
 #' @importFrom stats qchisq
-#' @return An S4 object of class \linkS4class{ConfregEL}.
+#' @return An object of class \linkS4class{ConfregEL}.
 #' @references Owen, Art. 1990. “Empirical Likelihood Ratio Confidence Regions.”
 #'   The Annals of Statistics 18 (1): 90–120. \doi{10.1214/aos/1176347494}.
 #' @seealso \link{confint}, \link{el_control}, \link{lht}, \link{plot}
@@ -127,7 +127,7 @@ setGeneric("confreg", function(object, parm, level = 0.95,
 #'
 #' @param object A fitted \linkS4class{EL} object.
 #' @param ... Not used.
-#' @return An S4 object of class \linkS4class{logLikEL}.
+#' @return An object of class \linkS4class{logLikEL}.
 #' @examples
 #' fit <- el_lm(formula = mpg ~ wt, data = mtcars)
 #' logLik(fit)

### 2022-05-03 R/el_glm.R

@@ -41,7 +41,7 @@
 #'   \deqn{H_{0j}: \beta_j = 0,\ j = 0, \dots, p-1.}
 #'   The test results are returned as \code{optim} and \code{parTests},
 #'   respectively.
-#' @return S4 object of class of \linkS4class{GLM}.
+#' @return An object of class of \linkS4class{GLM}.
 #' @references Chen, Song Xi, and Hengjian Cui. 2003.
 #'   “An Extended Empirical Likelihood for Generalized Linear Models.”
 #'   Statistica Sinica 13: 69–81.
@@ -150,7 +150,7 @@ el_glm <- function(formula, family = gaussian, data, weights = NULL, na.action,
   new("GLM",
     optim = el$optim, logp = el$logp, logl = el$logl, loglr = el$loglr,
     statistic = el$statistic, df = df, pval = pval, npar = p, weights = w,
-    dataMatrix = if (model) mm else matrix(NA_real_, nrow = 0L, ncol = 0L),
+    data = if (model) mm else matrix(NA_real_, nrow = 0L, ncol = 0L),
     coefficients = fit$coefficients, parTests = el$parTests,
     misc = list(
       family = fit$family, iter = fit$iter, converged = fit$converged,

### 2022-05-03 R/el_lm.R

@@ -44,7 +44,7 @@
 #'   \deqn{H_{0j}: \beta_j = 0,\ j = 0, \dots, p-1.}
 #'   The test results are returned as \code{optim} and \code{parTests},
 #'   respectively.
-#' @return An S4 object of class of \linkS4class{LM}.
+#' @return An object of class of \linkS4class{LM}.
 #' @references Owen, Art. 1991. “Empirical Likelihood for Linear Models.”
 #'   The Annals of Statistics 19 (4): 1725–47. \doi{10.1214/aos/1176348368}.
 #' @seealso \link{el_control}, \link{el_glm}, \link{lht}
@@ -82,10 +82,14 @@ el_lm <- function(formula, data, weights = NULL, na.action,
     x <- NULL
     mm <- cbind(y, x)
     return(new("LM",
-      optim = list(method = "lm", par = numeric(), lambda = numeric(),
-                   iterations = integer(), convergence = logical()),
-      misc = list(call = cl, terms = mt, xlevels = .getXlevels(mt, mf),
-                  na.action = attr(mf, "na.action"))
+      optim = list(
+        method = "lm", par = numeric(), lambda = numeric(),
+        iterations = integer(), convergence = logical()
+      ),
+      misc = list(
+        call = cl, terms = mt, xlevels = .getXlevels(mt, mf),
+        na.action = attr(mf, "na.action")
+      )
     ))
   } else {
     x <- model.matrix(mt, mf, NULL)
@@ -112,9 +116,11 @@ el_lm <- function(formula, data, weights = NULL, na.action,
   new("LM",
     optim = el$optim, logp = el$logp, logl = el$logl, loglr = el$loglr,
     statistic = el$statistic, df = df, pval = pval, npar = p, weights = w,
-    dataMatrix = if (model) mm else matrix(NA_real_, nrow = 0L, ncol = 0L),
+    data = if (model) mm else matrix(NA_real_, nrow = 0L, ncol = 0L),
     coefficients = z$coefficients, parTests = el$parTests,
-    misc = list(call = cl, terms = mt, xlevels = .getXlevels(mt, mf),
-                na.action = attr(mf, "na.action"))
+    misc = list(
+      call = cl, terms = mt, xlevels = .getXlevels(mt, mf),
+      na.action = attr(mf, "na.action")
+    )
   )
 }

### 2022-05-03 R/el_mean.R

@@ -11,11 +11,7 @@
 #' @param control A list of control parameters set by \code{\link{el_control}}.
 #' @param model A logical. If \code{TRUE} the data matrix used for model fitting
 #'   is returned.
-#' @return An S4 object of class \linkS4class{EL}.
-#' @references Glenn, N.L., and Yichuan Zhao. 2007.
-#'   “Weighted Empirical Likelihood Estimates and Their Robustness Properties.”
-#'   Computational Statistics & Data Analysis 51 (10): 5130–41.
-#'   \doi{10.1016/j.csda.2006.07.032}.
+#' @return An object of class \linkS4class{EL}.
 #' @references Owen, Art. 1990. “Empirical Likelihood Ratio Confidence Regions.”
 #'   The Annals of Statistics 18 (1): 90–120. \doi{10.1214/aos/1176347494}.
 #' @seealso \link{el_control}, \link{el_eval}, \link{lht}
@@ -73,7 +69,7 @@ el_mean <- function(par, x, weights = NULL, control = el_control(),
     statistic = el$statistic, df = p,
     pval = pchisq(el$statistic, df = p, lower.tail = FALSE), npar = p,
     weights = w,
-    dataMatrix = if (model) mm else matrix(NA_real_, nrow = 0L, ncol = 0L),
+    data = if (model) mm else matrix(NA_real_, nrow = 0L, ncol = 0L),
     coefficients = est
   )
 }

### 2022-05-03 R/eld-methods.R

@@ -6,16 +6,15 @@ setMethod(
     if (is(object, "GLM")) {
       stop("'eld' method is not applicable to a 'GLM' object")
     }
-    if (length(object@dataMatrix) == 0L) {
-      stop("'object' has no 'dataMatrix'; fit the model with 'model' = TRUE")
+    if (length(object@data) == 0L) {
+      stop("'object' has no 'data'; fit the model with 'model' = TRUE")
     }
     if (!is(control, "ControlEL")) {
       stop("invalid 'control' specified")
     }
     new("ELD", eld = eld_(
-      object@optim$method, object@coefficients, object@dataMatrix,
-      control@maxit_l, control@tol_l, control@th, control@nthreads,
-      object@weights
+      object@optim$method, object@coefficients, object@data, control@maxit_l,
+      control@tol_l, control@th, control@nthreads, object@weights
     ))
   }
 )

### 2022-05-03 R/lht.R

@@ -14,11 +14,10 @@
 #'   See ‘Details’.
 #' @param control A list of control parameters set by \code{\link{el_control}}.
 #' @details \code{\link{lht}} performs the constrained minimization of
-#'   \eqn{l(\theta)} introduced in \code{\link{el_control}}.
-#'   \code{rhs} and \code{lhs} cannot be both \code{NULL}. For non-\code{NULL}
-#'   \code{lhs}, it is required that \code{lhs} have full row rank
-#'   \eqn{q \leq p} and \eqn{p} be equal to \code{object$npar}, the number of
-#'   parameters in the fitted model.
+#'   \eqn{l(\theta)} described in \linkS4class{CEL}. \code{rhs} and \code{lhs}
+#'   cannot be both \code{NULL}. For non-\code{NULL} \code{lhs}, it is required
+#'   that \code{lhs} have full row rank \eqn{q \leq p} and \eqn{p} be equal to
+#'   \code{object$npar}, the number of parameters in the fitted model.
 #'
 #'   Depending on the specification of \code{rhs} and \code{lhs}, we have the
 #'   following three cases:
@@ -33,11 +32,12 @@
 #'   and the problem reduces to evaluating at \eqn{r} as
 #'   \deqn{l(r).}
 #'   }
-#' @return An S4 object of class of class \linkS4class{EL} or
-#'   \linkS4class{MinEL}.
-#' @references Kim, E., MacEachern, S., and Peruggia, M., (2021),
-#' "Empirical Likelihood for the Analysis of Experimental Designs,"
-#' \href{https://arxiv.org/abs/2112.09206}{arxiv:2112.09206}.
+#' @return If \code{lhs} is \code{NULL}, an object of class \linkS4class{EL}
+#' is returned. Otherwise, an object of class \linkS4class{CEL} is returned.
+#' @references Adimari, Gianfranco, and Annamaria Guolo. 2010.
+#'   “A Note on the Asymptotic Behaviour of Empirical Likelihood Statistics.”
+#'   Statistical Methods & Applications 19 (4): 463–76.
+#'   \doi{10.1007/s10260-010-0137-9}.
 #' @references Qin, Jing, and Jerry Lawless. 1995.
 #'   “Estimating Equations, Empirical Likelihood and Constraints on Parameters.”
 #'   Canadian Journal of Statistics 23 (2): 145–59. \doi{10.2307/3315441}.
[…]

### 2022-05-03 tests/testthat/test-eld.R

@@ -14,7 +14,7 @@ test_that("invalid 'object", {
   lhs <- matrix(c(1, -1), nrow = 1)
   fit2 <- lht(fit, lhs = lhs)
   expect_error(eld(df))
-  fit@dataMatrix <- matrix(NA_real_, nrow = 0L, ncol = 0L)
+  fit@data <- matrix(NA_real_, nrow = 0L, ncol = 0L)
   expect_error(eld(fit))
   expect_error(eld(fit2))
 })

### 2022-06-14 R/eld-methods.R
v1.5.2 release check.

@@ -13,7 +13,7 @@ setMethod(
       stop("invalid 'control' specified")
     }
     new("ELD", eld = eld_(
-      getMethod(object), coef(object), object@data, control@maxit_l,
+      getMethodEL(object), coef(object), object@data, control@maxit_l,
       control@tol_l, control@th, control@nthreads, object@weights
     ))
   }

### 2022-06-14 R/lht.R

@@ -4,14 +4,13 @@
 #'   \linkS4class{EL}.
 #'
 #' @param object A fitted \linkS4class{EL} object.
-#' @param rhs A numeric vector for the right-hand-side of hypothesis, with as
-#'   many entries as the rows in \code{lhs}. Defaults to \code{NULL}. See
-#'   ‘Details’.
-#' @param lhs A numeric matrix, or an object that can be coerced to a numeric
-#'   matrix. It specifies the left-hand-side of hypothesis. Each row gives a
-#'   linear combination of parameters. The number of columns should be equal to
-#'   the number of parameters in \code{object}. Defaults to \code{NULL}.
-#'   See ‘Details’.
+#' @param rhs A numeric vector or a column matrix for the right-hand-side of
+#'   hypothesis, with as many entries as the rows in \code{lhs}. Defaults to
+#'   \code{NULL}. See ‘Details’.
+#' @param lhs A numeric matrix or a vector (treated as a row matrix) for the
+#'   left-hand-side of hypothesis. Each row gives a linear combination of the
+#'   parameters in \code{object}. The number of columns should be equal to the
+#'   number of parameters. Defaults to \code{NULL}. See ‘Details’.
 #' @param control A list of control parameters set by \code{\link{el_control}}.
 #' @details \code{\link{lht}} performs the constrained minimization of
 #'   \eqn{l(\theta)} described in \linkS4class{CEL}. \code{rhs} and \code{lhs}
@@ -25,9 +24,9 @@
 #'   \item If both \code{rhs} and \code{lhs} are non-\code{NULL}, the
 #'   constrained minimization is performed with the right-hand-side \eqn{r} and
 #'   the left-hand-side \eqn{L} as
-#'   \deqn{\min_{\theta: L\theta = r} l(\theta).}
+#'   \deqn{\inf_{\theta: L\theta = r} l(\theta).}
 #'   \item If \code{rhs} is \code{NULL}, \eqn{r} is set to the zero vector as
-#'   \deqn{\min_{\theta: L\theta = 0} l(\theta).}
+#'   \deqn{\inf_{\theta: L\theta = 0} l(\theta).}
 #'   \item If \code{lhs} is \code{NULL}, \eqn{L} is set to the identity matrix
 #'   and the problem reduces to evaluating at \eqn{r} as
[…]

### 2022-06-14 tests/testthat/test-lht.R

@@ -82,11 +82,23 @@ test_that("when lht == eval", {
   expect_equal(fit@optim$lambda, fit2@optim$lambda)
 })
 
-test_that("invalid 'control", {
+test_that("invalid 'control'", {
   skip_on_os("windows", arch = "i386")
   n <- 10
   x <- rnorm(n)
   par <- runif(1, min(x), max(x))
   fit <- el_mean(par, x)
   expect_error(lht(fit, lhs = 1, control = list(maxit = 200L)))
+})
+
+test_that("vector lhs", {
+  n <- 100
+  x <- rnorm(n)
+  x2 <- rnorm(n)
+  y <- 1 + x + x2 + rnorm(n)
+  df <- data.frame(y, x, x2)
+  optcfg <- el_control(tol = 1e-08, th = 1e+10)
+  fit <- el_lm(y ~ x + x2, df, control = optcfg)
+  expect_error(lht(fit, lhs = c(1, -1, 0, 0)))
+  expect_s4_class(lht(fit, lhs = c(1, -1, 0)), "CEL")
 })

### 2022-06-17 README.Rmd
v1.5.2 release check.
cv argument defaults to NULL.

@@
 The R package **melt** provides a unified framework for data analysis with empirical likelihood methods.
-A collection of functions are available for basic regression analysis and hypothesis testing.
+A collection of functions are available for regression analysis and hypothesis testing.
 Much of its functionality and syntax mimics the corresponding base R functions.
@@
                 0, 0, 1, -1), byrow = TRUE, nrow = 3)
-lht(fit3, lhs = lhs)
+elt(fit3, lhs = lhs)
 

### 2022-06-17 R/checker.R

@@ -100,20 +100,18 @@ check_hypothesis <- function(lhs, rhs, p) {
   list(l = lhs, r = rhs)
 }
 
-check_cv <- function(cv, th) {
-  # cv <- tryCatch(as.numeric(cv),
-  #   warning = function(w) NA,
-  #   error = function(e) NA
-  # )
-  # if (any(length(cv) != 1L, is.na(cv), is.infinite(cv))) {
-  #   stop("'cv' is not a number")
-  # }
-  # if (cv < .Machine$double.eps) {
-  #   stop("'cv' is too small")
-  # }
-  # if (!is.null(th) && cv > 2 * th) {
-  #   stop("'cv' is too large")
-  # }
+check_level_ <- function(level) {
+  stopifnot(
+    "'level' is not a single numeric" = (is.numeric(level)),
+    "'level' is not a single numeric" = (length(level) == 1L),
+    "'level' is not a finite single numeric" = (is.finite(level)),
+    "'level' must be a single numeric between 0 and 1" =
+      (isTRUE(level >= 0 && level <= 1))
+  )
+  level
+}
+
+check_cv_ <- function(cv, th) {
   stopifnot(
     "'cv' is not a single numeric2" = (is.numeric(cv)),
     "'cv' is not a single numeric" = (length(cv) == 1L),

### 2022-06-17 R/el_eval.R

@@ -37,37 +37,29 @@
 #'   The Annals of Statistics 22 (1): 300–325. \doi{10.1214/aos/1176325370}.
 #' @seealso \link{el_control}
 #' @examples
-#' # test for variance with known mean
-#' x <- rnorm(100L)
-#' sigma <- 1
-#' g <- x^2 - sigma^2
-#' el_eval(g)
+#' x <- rnorm(50)
+#' par <- 0
+#' g <- x - par
+#' el_eval(g, weights = rep(c(1, 2), each = 25))
 #' @importFrom stats pchisq
 #' @export
 el_eval <- function(g, weights = NULL, control = el_control()) {
   mm <- as.matrix(g)
   n <- nrow(mm)
   p <- ncol(mm)
-  if (!is.numeric(mm) || !all(is.finite(mm))) {
-    stop("'g' must be a finite numeric matrix")
-  }
-  if (n < 2L) {
-    stop("not enough 'g' observations")
-  }
-  if (get_rank_(mm) != p || n <= p) {
-    stop("'g' must have full column rank")
-  }
-  if (!is(control, "ControlEL")) {
-    stop("invalid 'control' specified")
-  }
+  stopifnot(
+    "not enough observations in 'g' " = (n >= 2L),
+    "'g' must have full column rank" = (n > p),
+    "'g' must be a numeric matrix" = (is.numeric(mm)),
+    "'g' must be a finite numeric matrix" = (all(is.finite(mm))),
+    "'g' must have full column rank" = (get_rank_(mm) == p),
+    "invalid 'control' specified" = (is(control, "ControlEL"))
+  )
   w <- check_weights(weights, n)
   out <- eval_g_(mm, control@maxit_l, control@tol_l, control@th, w)
   out$df <- p
-  out$p.value <- pchisq(out$statistic, df = out$df, lower.tail = FALSE)
+  out$p.value <- pchisq(q = out$statistic, df = out$df, lower.tail = FALSE)
   out$npar <- p
-  # if (!is.null(weights)) {
-  #   out$weights <- w
-  # }
   out$weights <- w
   out
 }

### 2022-06-17 R/el_glm.R

@@ -46,7 +46,7 @@
 #' @references Chen, Song Xi, and Hengjian Cui. 2003.
 #'   “An Extended Empirical Likelihood for Generalized Linear Models.”
 #'   Statistica Sinica 13: 69–81.
-#' @seealso \link{el_control}, \link{el_lm}, \link{lht}
+#' @seealso \link{el_control}, \link{el_lm}, \link{elt}
 #' @examples
 #' n <- 50
 #' x <- rnorm(n)
@@ -222,7 +222,7 @@ el_glm <- function(formula, family = gaussian, data, weights = NULL, na.action,
 #' #' @references Chen, Song Xi, and Hengjian Cui. 2003.
 #' #'   “An Extended Empirical Likelihood for Generalized Linear Models.”
 #' #'   Statistica Sinica 13: 69–81.
-#' #' @seealso \link{el_control}, \link{el_lm}, \link{lht}
+#' #' @seealso \link{el_control}, \link{el_lm}, \link{elt}
 #' #' @examples
 #' #' n <- 50
 #' #' x <- rnorm(n)

### 2022-06-17 tests/testthat/test-el_mean.R

@@ -110,11 +110,11 @@ test_that("identical results for repeated executions", {
   expect_equal(wfit1, wfit2)
 
   lhs <- c(1, 0)
-  lht1 <- lht(fit1, lhs = lhs)
-  lht2 <- lht(fit2, lhs = lhs)
-  expect_equal(lht1, lht2)
+  elt1 <- elt(fit1, lhs = lhs)
+  elt2 <- elt(fit2, lhs = lhs)
+  expect_equal(elt1, elt2)
 
-  lht3 <- lht(wfit1, lhs = lhs)
-  lht4 <- lht(wfit2, lhs = lhs)
-  expect_equal(lht3, lht4)
+  elt3 <- elt(wfit1, lhs = lhs)
+  elt4 <- elt(wfit2, lhs = lhs)
+  expect_equal(elt3, elt4)
 })

### 2022-06-17 tests/testthat/test-eld.R

@@ -12,7 +12,7 @@ test_that("invalid 'object", {
   plot(eld)
   expect_length(eld@eld, n)
   lhs <- matrix(c(1, -1), nrow = 1)
-  fit2 <- lht(fit, lhs = lhs)
+  fit2 <- elt(fit, lhs = lhs)
   expect_error(eld(df))
   fit@data <- matrix(NA_real_, nrow = 0L, ncol = 0L)
   expect_error(eld(fit))

### 2022-06-21 README.Rmd
el_mean() arguments order changed.

@@
 <!-- badges: start -->
+[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
 [![CRAN status](https://www.r-pkg.org/badges/version/melt)](https://CRAN.R-project.org/package=melt)

### 2022-06-21 R/confreg-methods.R

@@ -1,74 +1,75 @@
 #' @rdname confreg
-setMethod(
-  "confreg", "EL",
-  function(object, parm, level = 0.95, cv = NULL, npoints = 50L,
-           control = el_control()) {
-    est <- coef(object)
-    if (length(est) == 0L) {
-      stop("'confreg' method is not applicable to an empty model")
-    } else if (length(est) == 1L) {
-      stop("'confreg' method is not applicable to a model with one parameter")
+setMethod("confreg", "EL", function(object,
+                                    parm,
+                                    level = 0.95,
+                                    cv = NULL,
+                                    npoints = 50L,
+                                    control = el_control()) {
+  est <- coef(object)
+  if (length(est) == 0L) {
+    stop("'confreg' method is not applicable to an empty model")
+  } else if (length(est) == 1L) {
+    stop("'confreg' method is not applicable to a model with one parameter")
+  }
+  pnames <- if (is.null(names(est))) seq(length(est)) else names(est)
+  if (!missing(parm)) {
+    if (length(parm) != 2L) {
+      stop("length of 'parm' must be two")
     }
-    pnames <- if (is.null(names(est))) seq(length(est)) else names(est)
-    if (!missing(parm)) {
-      if (length(parm) != 2L) {
-        stop("length of 'parm' must be two")
-      }
-      if (is.numeric(parm) && all(is.finite(parm))) {
-        idx <- as.integer(parm)
-        est <- est[idx]
-        pnames <- pnames[idx]
-        if (length(unique(est)) != 2L) {
-          stop("only one parameter specified by 'parm'")
-        }
-      } else if (is.character(parm)) {
-        if (is.null(names(est))) {
-          stop(
-            "'parm' is not recognized since 'object' has no named parameters"
-          )
-        }
-        idx <- match(parm, pnames)
-        est <- est[idx]
-        pnames <- pnames[idx]
-        if (length(unique(est)) != 2L) {
-          stop("only one parameter specified by 'parm'")
-        }
-      } else {
-        stop("invalid 'parm' specified")
-      }
-    } else {
-      idx <- c(1L, 2L)
+    if (is.numeric(parm) && all(is.finite(parm))) {
+      idx <- as.integer(parm)
       est <- est[idx]
[…]

### 2022-06-21 R/el_lm.R

@@ -63,8 +63,13 @@
 #' @importFrom stats .getXlevels is.empty.model lm.fit lm.wfit model.matrix
 #'   model.response model.weights pchisq
 #' @export
-el_lm <- function(formula, data, weights = NULL, na.action,
-                  control = el_control(), model = TRUE, ...) {
+el_lm <- function(formula,
+                  data,
+                  weights = NULL,
+                  na.action,
+                  control = el_control(),
+                  model = TRUE,
+                  ...) {
   if (!is(control, "ControlEL")) {
     stop("invalid 'control' specified")
   }

### 2022-06-21 R/el_mean.R

@@ -18,24 +18,27 @@
 #' @seealso \link{el_control}, \link{el_eval}, \link{elt}
 #' @examples
 #' # scalar mean
-#' par <- 0
 #' x <- rnorm(100)
-#' el_mean(par, x)
+#' par <- 0
+#' el_mean(x, par)
 #'
 #' # vector mean
-#' par <- c(0, 0)
 #' x <- matrix(rnorm(100), ncol = 2)
-#' el_mean(par, x)
+#' par <- c(0, 0)
+#' el_mean(x, par)
 #'
 #' # weighted data
-#' par <- c(0, 0)
 #' x <- matrix(rnorm(100), ncol = 2)
+#' par <- c(0, 0)
 #' w <- rep(c(1, 2), each = 25)
-#' el_mean(par, x, w)
+#' el_mean(x, par, w)
 #' @importFrom methods is new
 #' @importFrom stats pchisq
 #' @export
-el_mean <- function(par, x, weights = NULL, control = el_control(),
+el_mean <- function(x,
+                    par,
+                    weights = NULL,
+                    control = el_control(),
                     model = TRUE) {
   mm <- as.matrix(x)
   n <- nrow(mm)

### 2022-06-21 R/eld-methods.R

@@ -1,18 +1,15 @@
 #' @rdname eld
 #' @importFrom methods is
-setMethod(
-  "eld", "EL",
-  function(object, control = el_control()) {
-    stopifnot(
-      "'eld' method is not applicable to a 'GLM' object" =
-        (!is(object, "GLM")),
-      "'object' has no 'data'; fit the model with 'model' = TRUE" =
-        (length(getDataMatrix(object)) > 1L),
-      "invalid 'control' specified" = (is(control, "ControlEL"))
-    )
-    new("ELD", eld = eld_(
-      getMethodEL(object), coef(object), getDataMatrix(object), control@maxit_l,
-      control@tol_l, control@th, control@nthreads, object@weights
-    ))
-  }
-)
+setMethod("eld", "EL", function(object, control = el_control()) {
+  stopifnot(
+    "'eld' method is not applicable to a 'GLM' object" =
+      (!is(object, "GLM")),
+    "'object' has no 'data'; fit the model with 'model' = TRUE" =
+      (length(getDataMatrix(object)) > 1L),
+    "invalid 'control' specified" = (is(control, "ControlEL"))
+  )
+  new("ELD", eld = eld_(
+    getMethodEL(object), coef(object), getDataMatrix(object), control@maxit_l,
+    control@tol_l, control@th, control@nthreads, object@weights
+  ))
+})

### 2022-06-21 R/elmt.R

@@ -3,6 +3,29 @@
 #' #' Tests a linear hypothesis for objects that inherit from class
 #' #'   \linkS4class{EL}.
 #' #'
+#' #' @export
+#' elmt <- function() {
+#'   cat("multiple test")
+#' }
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+#' #' Multiple tests with empirical likelihood
+#' #'
+#' #' Tests a linear hypothesis for objects that inherit from class
+#' #'   \linkS4class{EL}.
+#' #'
 #' #' @param object A fitted \linkS4class{EL} object.
 #' #' @param rhs A numeric vector for the right-hand-side of hypothesis, with as
 #' #'   many entries as the rows in \code{lhs}. Defaults to \code{NULL}. See

### 2022-06-21 R/elt.R

@@ -76,8 +76,12 @@
 #' @importFrom methods is
 #' @importFrom stats pchisq
 #' @export
-elt <- function(object, rhs = NULL, lhs = NULL, alpha = 0.05,
-                calibrate = c("chisq", "boot", "f"), control = el_control()) {
+elt <- function(object,
+                rhs = NULL,
+                lhs = NULL,
+                alpha = 0.05,
+                calibrate = c("chisq", "boot", "f"),
+                control = el_control()) {
   stopifnot(
     "invalid 'object' supplied" = (is(object, "EL")),
     "'object' has no 'data'; fit the model with 'model' = TRUE" =

### 2022-06-21 R/summary-methods.R

@@ -1,31 +1,28 @@
 #' @describeIn summary Summarizes the results of the overall test and the
 #'   tests for each parameter.
 #' @importFrom stats pchisq
-setMethod(
-  "summary", "LM",
-  function(object, ...) {
-    z <- object
-    p <- z@npar
-    if (p == 0L) {
-      return(new("SummaryLM",
-        statistic = z@statistic, df = z@df, convergence = z@optim$convergence,
-        parMatrix = matrix(NA_real_, 0L, 3L,
-          dimnames = list(NULL, c("Estimate", "Chisq", "Pr(>Chis)"))
-        ),
-        weighted = length(z@weights) != 0L, na.action = z@misc$na.action,
-        call = z@misc$call, terms = z@misc$terms,
-        aliased = is.na(coef(z))
-      ))
-    }
-    new("SummaryLM",
+setMethod("summary", "LM", function(object, ...) {
+  z <- object
+  p <- z@npar
+  if (p == 0L) {
+    return(new("SummaryLM",
       statistic = z@statistic, df = z@df, convergence = z@optim$convergence,
-      parMatrix = cbind(
-        Estimate = coef(z),
-        Chisq = z@parTests$statistic,
-        `Pr(>Chisq)` = pchisq(z@parTests$statistic, df = 1L, lower.tail = FALSE)
+      parMatrix = matrix(NA_real_, 0L, 3L,
+        dimnames = list(NULL, c("Estimate", "Chisq", "Pr(>Chis)"))
       ),
       weighted = length(z@weights) != 0L, na.action = z@misc$na.action,
-      call = z@misc$call, terms = z@misc$terms, aliased = is.na(coef(z))
-    )
+      call = z@misc$call, terms = z@misc$terms,
+      aliased = is.na(coef(z))
+    ))
   }
-)
+  new("SummaryLM",
+    statistic = z@statistic, df = z@df, convergence = z@optim$convergence,
+    parMatrix = cbind(
+      Estimate = coef(z),
+      Chisq = z@parTests$statistic,
+      `Pr(>Chisq)` = pchisq(z@parTests$statistic, df = 1L, lower.tail = FALSE)
+    ),
+    weighted = length(z@weights) != 0L, na.action = z@misc$na.action,
+    call = z@misc$call, terms = z@misc$terms, aliased = is.na(coef(z))
+  )
+})

### 2022-06-21 tests/testthat/test-logLik.R

@@ -4,7 +4,7 @@ test_that("logLik at maximum EL estimates", {
   x <- rnorm(n)
   par <- runif(1, min(x), max(x))
   optcfg <- el_control(maxit_l = 200L, tol_l = 1e-08, th = 1e+10)
-  fit <- el_mean(par, x, control = optcfg)
+  fit <- el_mean(x, par, control = optcfg)
   loglik <- suppressWarnings(logLik(fit, REML = T))
   expect_output(show(loglik))
   expect_output(print(loglik))

### 2022-06-23 README.Rmd
badge updated

--- 2022-06-17 README.Rmd
@@
 <!-- badges: start -->
+[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
 [![CRAN status](https://www.r-pkg.org/badges/version/melt)](https://CRAN.R-project.org/package=melt)
-[![R-CMD-check](https://github.com/markean/melt/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/markean/melt/actions/workflows/R-CMD-check.yaml)
+[![R-CMD-check](https://github.com/markean/melt/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/markean/melt/actions/workflows/check-standard.yaml)
 [![Codecov test coverage](https://codecov.io/gh/markean/melt/branch/master/graph/badge.svg)](https://app.codecov.io/gh/markean/melt?branch=master)

### 2022-06-28 README.Rmd
unit tests improved

@@
 Additional functions are available for multiple testing for the analysis of experimental designs.
-Details of the testing procedures are given in [Kim et al.
-(2021)](https://arxiv.org/abs/2112.09206).
+Details of the testing procedures are given in [Kim, MacEachern, and Peruggia (2021)](https://arxiv.org/abs/2112.09206).
 
@@
 # one sample test for mean
-fit1 <- el_mean(par = 0, x = rnorm(n = 100))
+fit1 <- el_mean(rnorm(n = 100), par = 0)
 confint(fit1)
@@
 # linear regression
-fit2 <- el_lm(formula = mpg ~ wt, data = mtcars)
+fit2 <- el_lm(mpg ~ wt, data = mtcars)
 summary(fit2)
@@
 data("clothianidin")
-fit3 <- el_lm(clo ~ -1 + trt, clothianidin)
+fit3 <- el_lm(clo ~ -1 + trt, data = clothianidin)
 summary(fit3)

### 2022-06-28 R/AllClasses.R

@@ -207,7 +207,8 @@ setClass("ConfregEL",
 #' @slot nthreads A single integer for the number of threads for parallel
 #'   computation via OpenMP (if available).
 #' @slot seed A single integer for the seed for random number generation.
-#' @slot B A single integer for the number of bootstrap replicates.
+#' @slot b A single integer for the number of bootstrap replicates.
+#' @slot m A single integer for the number of Monte Carlo samples.
 #' @seealso \link{el_control}
 #' @examples
 #' showClass("ControlEL")
@@ -215,7 +216,7 @@ setClass("ControlEL",
   slots = c(
     maxit = "integer", maxit_l = "integer", tol = "numeric", tol_l = "numeric",
     step = "ANY", th = "ANY", nthreads = "integer", seed = "integer",
-    B = "integer"
+    b = "integer", m = "integer"
   )
 )
 
@@ -254,16 +255,20 @@ setClass("ELT",
 
 #' S4 class \linkS4class{ELMT}
 #'
-#' S4 class for multiple empirical likelihood test.
+#' S4 class for empirical likelihood multiple tests.
 #'
-#' @slot statistic A single numeric for the minus twice the (constrained)
-#'   empirical log-likelihood ratio.
+#' @slot alpha A single numeric for the overall significance level.
+#' @slot statistic A numeric vector for the minus twice the (constrained)
+#'   empirical log-likelihood ratios.
+#' @slot cv A single numeric for the multiplicity adjusted critical value.
+#' @slot pval A numeric vector for the multiplicity adjusted \eqn{p}-values.
+#' @slot calibrate A single character for the calibration method used.
 #' @examples
 #' showClass("ELMT")
 setClass("ELMT",
   slots = c(
-    alpha = "numeric", statistic = "numeric", cv = "numeric",
-    pval = "numeric", calibrate = "character"
+    alpha = "numeric", statistic = "numeric", cv = "numeric", pval = "numeric",
+    calibrate = "character"
   )
 )
 

### 2022-06-28 R/calibrate.R

@@ -23,7 +23,7 @@ calibrate <- function(calibrate, alpha, statistic, p, par, object, control) {
     },
     "boot" = {
       out <- computeBootstrapStatistics(
-        control@B, control@seed, control@nthreads, getMethodEL(object),
+        control@b, control@seed, control@nthreads, getMethodEL(object),
         getDataMatrix(object), par, control@maxit_l, control@tol_l, control@th,
         getWeights(object)
       )

### 2022-06-28 tests/testthat/test-el_lm.R

@@ -52,7 +52,7 @@ test_that("empty model", {
 })
 
 test_that("no intercept", {
-  fit <- el_lm(mpg ~ -1 + ., mtcars)
+  fit <- el_lm(mpg ~ -1 + ., data = mtcars)
   expect_s4_class(fit, "LM")
 })
 

### 2022-06-28 tests/testthat/test-elt.R

@@ -31,7 +31,7 @@ test_that("invalid 'calibrate'", {
 })
 
 test_that("invalid 'control'", {
-  fit <- el_mean(sleep$extra, 0)
+  fit <- el_mean(sleep$extra, par = 0)
   expect_error(elt(fit, lhs = 1, control = list(maxit = 200L)))
 })
 

### 2022-06-30 src/testMultipleHypotheses.cpp
vignette added

@@ -24,6 +24,7 @@ Rcpp::List testMultipleHypotheses(const double alpha,
                                   const Rcpp::Nullable<double> th,
                                   const Eigen::Map<Eigen::ArrayXd>& w) {
   const double gamma = setStep(x.rows(), step);
+  // 1. compute test statistics
   std::vector<double> test_statistic(m);
   for (int j = 0; j < m; ++j) {
     const Eigen::VectorXd r = rhs.middleRows(q(j), q(j + 1) - q(j));
@@ -33,15 +34,12 @@ Rcpp::List testMultipleHypotheses(const double alpha,
                  test_th, w);
     test_statistic[j] = 2.0 * el.nllr;
   }
-
-
-
+  // 2. compute cutoff value
   // sample covariance with plug-in estimate
   const Eigen::MatrixXd s = cov(method, est, x);
   // sqrt of sample covariance
   const Eigen::SelfAdjointEigenSolver<Eigen::MatrixXd> es(s);
   const Eigen::MatrixXd sqrt_s = es.operatorSqrt();
-
   // ahat matrices
   const int p = est.size();
   const Eigen::MatrixXd h = dg0_inv(method, x);
@@ -50,8 +48,7 @@ Rcpp::List testMultipleHypotheses(const double alpha,
     const Eigen::MatrixXd l = lhs.middleRows(q(j), q(j + 1) - q(j));
     amat.middleCols(j * p, p) = ahat(l, h, s);
   }
-
-  // 4. random monte carlo
+  // Monte Carlo simulation
   std::vector<double> max_statistic(M);
   // #pragma omp parallel for num_threads(nthreads)
   for (int b = 0; b < M; ++b) {
@@ -61,8 +58,7 @@ Rcpp::List testMultipleHypotheses(const double alpha,
     max_statistic[b] = (u * tmp).maxCoeff();
   }
   const double cv = computeQuantile(Rcpp::wrap(max_statistic), 1 - alpha);
-
-  // adjusted p-values
+  // 3. compute adjusted p-values
   std::vector<double> adj_pval(m);
   for (int j = 0; j < m; ++j) {
     adj_pval[j] =
@@ -71,8 +67,6 @@ Rcpp::List testMultipleHypotheses(const double alpha,
         [test_statistic, j](double x) {return (x > test_statistic[j]);}) /
           static_cast<double>(M);
   }
-
-  // critical value can be computed independent of the statistics
[…]

### 2022-06-30 R/confreg-methods.R
pkgdown added

@@ -5,6 +5,10 @@ setMethod("confreg", "EL", function(object,
                                     cv = NULL,
                                     npoints = 50L,
                                     control = el_control()) {
+  if (npoints > 3000L) {
+    stop("sefsef")
+  }
+
   if (!is(control, "ControlEL")) {
     stop("invalid `control` specified.")
   }

### 2022-06-30 R/elt-methods.R
temporary fix for missing doc

@@ -59,7 +59,7 @@ setMethod("elt", "EL", function(object,
   )
 })
 
-
+#' @rdname elt
 setMethod("elt", "missing",function(object,
                                     rhs = NULL,
                                     lhs = NULL,

### 2022-07-02 R/AllClasses.R
test srrstats tags

@@ -66,6 +66,11 @@
 #'   The Annals of Statistics 22 (1): 300–325. \doi{10.1214/aos/1176325370}.
 #' @examples
 #' showClass("EL")
+#' @srrstats {G1.0} References are included throughout the package
+#'   documentation.
+#' @srrstats {G1.4} `roxygen2` is used to document all functions.
+#' @srrstats {G1.3} Statistical terminology is unambiguously defined throughout
+#'   the package documentation.
 setClass("EL",
   slots = c(
     optim = "list", logp = "numeric", logl = "numeric", loglr = "numeric",
@@ -129,6 +134,11 @@ setClass("EL",
 #'   Canadian Journal of Statistics 23 (2): 145–59. \doi{10.2307/3315441}.
 #' @examples
 #' showClass("CEL")
+#' @srrstats {G1.1} The package attempts the first implementation of the nested
+#'   bilevel optimization approach within R to compute constrained empirical
+#'   likelihood. The inner layer Newton-Raphson method for empirical likelihood
+#'   is written in C++, enabling faster computation than other packages written
+#'   in R.
 setClass("CEL", contains = "EL")
 
 

### 2022-07-02 R/confint-methods.R

@@ -1,4 +1,5 @@
 #' @rdname confint
+#' @srrstats {RE4.3} `confint()` method is exported.
 setMethod("confint", "EL", function(object,
                                     parm,
                                     level = 0.95,

### 2022-07-05 R/srr-stats-standards.R
some methods added

@@ -11,6 +11,11 @@
 #'
 #' @srrstats {G1.2} Life cycle statement is included in
 #'   `.github/CONTRIBUTING.md`.
+#' @srrstats {G2.15} All user-facing functions in the package inspect missing
+#'   values in inputs. Some functions produce an error and some functions allow
+#'   users to apply appropriate treatments to the missing values. These inputs
+#'   are not passed to base routines like `mean()`, `sd()`, `cor()`, etc. See
+#'   `R/validate.R` as well.
 #' @srrstats {G5.2, G5.2a, G5.2b} Every message produced within the package is
 #'   unique. Unit tests in `test/testthat` cover all the conditions that trigger
 #'   any message. See `R/validate.R` as well.
@@ -29,11 +34,6 @@
 #' @srrstatsTODO {G2.11} *Software should ensure that `data.frame`-like tabular objects which have columns which do not themselves have standard class attributes (typically, `vector`) are appropriately processed, and do not error without reason. This behaviour should be tested. Again, columns created by the [`units` package](https://github.com/r-quantities/units/) provide a good test case.*
 
 #' @srrstatsTODO {G2.12} *Software should ensure that `data.frame`-like tabular objects which have list columns should ensure that those columns are appropriately pre-processed either through being removed, converted to equivalent vector columns where appropriate, or some other appropriate treatment such as an informative error. This behaviour should be tested.*
-
-
-
-#' @srrstatsTODO {G2.15} *Functions should never assume non-missingness, and should never pass data with potential missing values to any base routines with default `na.rm = FALSE`-type parameters (such as [`mean()`](https://stat.ethz.ch/R-manual/R-devel/library/base/html/mean.html), [`sd()`](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/sd.html) or [`cor()`](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/cor.html)).*
-
 
 
 
@@ -58,17 +58,10 @@
 #' @srrstatsTODO {RE2.3} *Where applicable, Regression Software should enable data to be centred (for example, through converting to zero-mean equivalent values; or to z-scores) or offset (for example, to zero-intercept equivalent values) via additional parameters, with the effects of any such parameters clearly documented and tested.*
 
 
-
[…]

### 2022-07-05 src/utils.cpp

@@ -7,315 +7,387 @@
 
 std::function<Eigen::MatrixXd(const Eigen::Ref<const Eigen::MatrixXd> &,
                               const Eigen::Ref<const Eigen::VectorXd> &)>
-  g_fn2(const std::string method) {
-    std::map<std::string, std::function<Eigen::MatrixXd(
-        const Eigen::Ref<const Eigen::MatrixXd> &,
-        const Eigen::Ref<const Eigen::VectorXd> &)>>
-          g_map{{{"mean", g_mean},
-          {"lm", g_lm},
-          {"gaussian_identity", g_lm},
-          {"gaussian_log", g_gauss_log},
-          {"gaussian_inverse", g_gauss_inverse},
-          {"binomial_logit", g_bin_logit},
-          {"binomial_probit", g_bin_probit},
-          {"binomial_log", g_bin_log},
-          {"poisson_log", g_poi_log},
-          {"poisson_identity", g_poi_identity},
-          {"poisson_sqrt", g_poi_sqrt}}};
-    return g_map[method];
+g_fn2(const std::string method)
+{
+  std::map<std::string, std::function<Eigen::MatrixXd(
+                            const Eigen::Ref<const Eigen::MatrixXd> &,
+                            const Eigen::Ref<const Eigen::VectorXd> &)>>
+      g_map{{{"mean", g_mean},
+             {"lm", g_lm},
+             {"gaussian_identity", g_lm},
+             {"gaussian_log", g_gauss_log},
+             {"gaussian_inverse", g_gauss_inverse},
+             {"binomial_logit", g_bin_logit},
+             {"binomial_probit", g_bin_probit},
+             {"binomial_log", g_bin_log},
+             {"poisson_log", g_poi_log},
+             {"poisson_identity", g_poi_identity},
+             {"poisson_sqrt", g_poi_sqrt}}};
+  return g_map[method];
 }
 
-Eigen::ArrayXd inverse_linkinv(const Eigen::Ref<const Eigen::VectorXd> &x) {
+Eigen::ArrayXd inverse_linkinv(const Eigen::Ref<const Eigen::VectorXd> &x)
+{
   return x.array().inverse();
 }
-Eigen::ArrayXd log_linkinv(const Eigen::Ref<const Eigen::VectorXd> &x) {
+
[…]

### 2022-07-05 tests/testthat/test-el_lm.R

@@ -56,7 +56,7 @@ test_that("No intercept.", {
   expect_s4_class(fit, "LM")
 })
 
-test_that("Print method.", {
+test_that("`print()` method.", {
   fit <- el_lm(mpg ~ disp + hp, data = mtcars)
   expect_output(show(fit))
   expect_output(print(fit))
@@ -71,6 +71,12 @@ test_that("Print method.", {
   df2[1, 1] <- NA
   fit2 <- el_lm(mpg ~ disp + hp, data = df2)
   expect_output(print(summary(fit2)))
+})
+
+test_that("`formula()` and `nobs()` methods.", {
+  fit <- el_lm(mpg ~ disp + hp, data = mtcars)
+  expect_type(formula(fit), "language")
+  expect_identical(nobs(fit), nrow(mtcars))
 })
 
 #' @srrstats {G5.6, G5.6a, G5.6b} `el_lm()` returns the expected coefficients

### 2022-07-05 R/srr-stats-standards.R
adding vignette

@@ -44,24 +44,20 @@
 
 
 
-
-
-
-#' @srrstatsTODO {RE1.2} *Regression Software should document expected format (types or classes) for inputting predictor variables, including descriptions of types or classes which are not accepted.*
 #' @srrstatsTODO {RE1.3} *Regression Software which passes or otherwise transforms aspects of input data onto output structures should ensure that those output structures retain all relevant aspects of input data, notably including row and column names, and potentially information from other `attributes()`.*
 #' @srrstatsTODO {RE1.3a} *Where otherwise relevant information is not transferred, this should be explicitly documented.*
+
 
 #' @srrstatsTODO {RE2.0} *Regression Software should document any transformations applied to input data, for example conversion of label-values to `factor`, and should provide ways to explicitly avoid any default transformations (with error or warning conditions where appropriate).*
 
 
-
+#' offset!!
 #' @srrstatsTODO {RE2.3} *Where applicable, Regression Software should enable data to be centred (for example, through converting to zero-mean equivalent values; or to z-scores) or offset (for example, to zero-intercept equivalent values) via additional parameters, with the effects of any such parameters clearly documented and tested.*
 
 
 #' @srrstatsTODO {RE3.1} *Enable such messages to be optionally suppressed, yet should ensure that the resultant model object nevertheless includes sufficient data to identify lack of convergence.*
 
 
-#' @srrstatsTODO {RE4.1} *Regression Software may enable an ability to generate a model object without actually fitting values. This may be useful for controlling batch processing of computationally intensive fitting algorithms.*
 
 
 #' @srrstatsTODO {RE4.8} *Response variables, and associated "metadata" where applicable.*
@@ -72,7 +68,6 @@
 #' @srrstatsTODO {RE4.13} *Predictor variables, and associated "metadata" where applicable.*
 
 
-#' @srrstatsTODO {RE4.16} *Regression Software which models distinct responses for different categorical groups should include the ability to submit new groups to `predict()` methods.*
 
 
 #' @srrstatsTODO {RE5.0} *Scaling relationships between sizes of input data (numbers of observations, with potential extension to numbers of variables/columns) and speed of algorithm.*
@@ -117,15 +112,25 @@ NULL
 #'   We believe that the separate flags are not necessary for the package.
 #' @srrstatsNA {G5.11} Unit tests do not require large data sets.
[…]

### 2022-07-05 tests/testthat/test-el_lm.R

@@ -73,8 +73,9 @@ test_that("`print()` method.", {
   expect_output(print(summary(fit2)))
 })
 
-test_that("`formula()` and `nobs()` methods.", {
+test_that("`conv()`, `formula()`, and `nobs()` methods.", {
   fit <- el_lm(mpg ~ disp + hp, data = mtcars)
+  expect_false(conv(fit))
   expect_type(formula(fit), "language")
   expect_identical(nobs(fit), nrow(mtcars))
 })

### 2022-07-06 R/AllClasses.R
`keep_data` added in `el_control()`

@@ -79,13 +79,13 @@ setClass("EL",
   slots = c(
     optim = "list", logp = "numeric", logl = "numeric", loglr = "numeric",
     statistic = "numeric", df = "integer", pval = "numeric", nobs = "integer",
-    npar = "integer", weights = "numeric", data = "matrix",
+    npar = "integer", weights = "numeric", data = "ANY",
     coefficients = "numeric", method = "character"
   ),
   prototype = list(
     optim = list(), logp = numeric(), logl = numeric(), loglr = numeric(),
     statistic = numeric(), df = 0L, pval = numeric(), npar = 0L,
-    weights = numeric(), data = matrix(NA_real_, nrow = 0L, ncol = 0L),
+    weights = numeric(), data = NULL,
     coefficients = numeric(), method = NA_character_
   )
 )
@@ -228,6 +228,7 @@ setClass("ConfregEL",
 #'   log-likelihood ratio.
 #' @slot verbose A single logical for whether to print a message on the
 #'   convergence status.
+#' @slot keep_data A single logical for whether to
 #' @slot nthreads A single integer for the number of threads for parallel
 #'   computation via OpenMP (if available).
 #' @slot seed A single integer for the seed for random number generation.
@@ -240,8 +241,8 @@ setClass("ConfregEL",
 setClass("ControlEL",
   slots = c(
     maxit = "integer", maxit_l = "integer", tol = "numeric", tol_l = "numeric",
-    step = "ANY", th = "ANY", verbose = "logical", nthreads = "integer",
-    seed = "integer", b = "integer", m = "integer"
+    step = "ANY", th = "ANY", verbose = "logical", keep_data = "logical",
+    nthreads = "integer", seed = "integer", b = "integer", m = "integer"
   )
 )
 

### 2022-07-06 R/el_control.R

@@ -21,6 +21,9 @@
 #'   status is printed when fitting objects that inherit from class
 #'   \linkS4class{EL}.
 #'   Defaults to `FALSE`.
+#' @param keep_data A single logical. If `TRUE`, the data used for fitting
+#'   objects that inherit from class \linkS4class{EL} are kept in the `data`
+#'   slot for later use with other methods. Defaults to `TRUE`.
 #' @param nthreads A single integer for the number of threads for parallel
 #'   computation via OpenMP (if available). Defaults to the half of the
 #'   available threads. For better performance, it is generally recommended to
@@ -50,6 +53,10 @@
 #'   throughout the package documentation.
 #' @srrstats {RE3.2, RE3.3} Convergence thresholds can be set by the `tol` and
 #'   `tol_l` arguments with the default values documented.
+#' @srrstats {RE3.0, RE3.1} The `verbose` argument controls whether to print a
+#'   message on the convergence status when fitting objects. The status is
+#'   printed regardless of the `verbose` when `print()` method is used to the
+#'   fitted objects. Alternatively, `conv()` method also extracts the status.
 el_control <- function(maxit = 200L,
                        maxit_l = 25L,
                        tol = 1e-06,
@@ -57,6 +64,7 @@ el_control <- function(maxit = 200L,
                        step = NULL,
                        th = NULL,
                        verbose = FALSE,
+                       keep_data = TRUE,
                        nthreads,
                        seed = sample.int(.Machine$integer.max, 1L),
                        b = 10000L,
@@ -72,6 +80,7 @@ el_control <- function(maxit = 200L,
     th <- validate_th(th)
   }
   verbose <- validate_verbose(verbose)
+  keep_data <- validate_keep_data(keep_data)
   max_threads <- get_max_threads()
   if (missing(nthreads)) {
     nthreads <- as.integer(max(1L, max_threads / 2L))
@@ -83,6 +92,7 @@ el_control <- function(maxit = 200L,
   m <- validate_m(m)
   new("ControlEL",
     maxit = maxit, maxit_l = maxit_l, tol = tol, tol_l = tol_l, step = step,
-    th = th, verbose = verbose, nthreads = nthreads, seed = seed, b = b, m = m
+    th = th, verbose = verbose, keep_data = keep_data, nthreads = nthreads,
+    seed = seed, b = b, m = m
   )
 }

### 2022-07-06 R/el_mean.R

@@ -13,8 +13,6 @@
 #'   If non-`NULL`, weighted empirical likelihood is computed.
 #' @param control An object of class \linkS4class{ControlEL} constructed by
 #'   [el_control()].
-#' @param model A single logical. If `TRUE` the data matrix used for model
-#'   fitting is returned.
 #' @return An object of class \linkS4class{EL}.
 #' @references Owen A (1990).
 #'   “Empirical Likelihood Ratio Confidence Regions.”
@@ -51,8 +49,7 @@
 el_mean <- function(x,
                     par,
                     weights = NULL,
-                    control = el_control(),
-                    model = TRUE) {
+                    control = el_control()) {
   mm <- validate_x(x)
   stopifnot(
     "`par` must be a finite numeric vector." =
@@ -70,7 +67,6 @@ el_mean <- function(x,
   } else {
     est <- colMeans(mm)
   }
-  model <- validate_model(model)
   out <- compute_EL(
     "mean", par, mm, control@maxit_l, control@tol_l, control@th, w
   )
@@ -84,8 +80,7 @@ el_mean <- function(x,
     optim = out$optim, logp = out$logp, logl = out$logl, loglr = out$loglr,
     statistic = out$statistic, df = p,
     pval = pchisq(out$statistic, df = p, lower.tail = FALSE), nobs = n,
-    npar = p, weights = w,
-    data = if (model) mm else matrix(NA_real_, nrow = 0L, ncol = 0L),
+    npar = p, weights = w, data = if (control@keep_data) mm else NULL,
     coefficients = est, method = "mean"
   )
 }

### 2022-07-06 R/elmt-methods.R

@@ -6,9 +6,9 @@ setMethod("elmt", "EL", function(object,
                                  alpha = 0.05,
                                  control = el_control()) {
   stopifnot(
-    "`object` has no `data`; fit the model with `model` = TRUE." =
-      (length(getDataMatrix(object)) > 1L),
-    "invalid `control` specified." = (is(control, "ControlEL"))
+    "`object` has no `data`. Fit the model with `keep_data == TRUE`." =
+      (!is.null(getDataMatrix(object))),
+    "Invalid `control` specified." = (is(control, "ControlEL"))
   )
   h <- validate_hypotheses(rhs, lhs, object@npar)
   alpha <- validate_alpha(alpha)

### 2022-07-06 R/srr-stats-standards.R

@@ -53,11 +53,6 @@
 #' @srrstatsTODO {RE2.3} *Where applicable, Regression Software should enable data to be centred (for example, through converting to zero-mean equivalent values; or to z-scores) or offset (for example, to zero-intercept equivalent values) via additional parameters, with the effects of any such parameters clearly documented and tested.*
 
 
-#' @srrstatsTODO {RE3.1} *Enable such messages to be optionally suppressed, yet should ensure that the resultant model object nevertheless includes sufficient data to identify lack of convergence.*
-
-
-
-
 #' @srrstatsTODO {RE4.8} *Response variables, and associated "metadata" where applicable.*
 #' @srrstatsTODO {RE4.9} *Modelled values of response variables.*
 #' @srrstatsTODO {RE4.10} *Model Residuals, including sufficient documentation to enable interpretation of residuals, and to enable users to submit residuals to their own tests.*

### 2022-07-06 R/validate.R

@@ -128,6 +128,8 @@ validate_th <- function(th) {
 #'
 #' @param verbose A single logical.
 #' @return A single logical.
+#' @srrstats {G3.0} The package does not compare floating point numbers for
+#'   equality. All numeric equality comparisons are made between integers.
 #' @noRd
 validate_verbose <- function(verbose) {
   stopifnot(
@@ -135,6 +137,21 @@ validate_verbose <- function(verbose) {
       (isTRUE(is.logical(verbose) && length(verbose) == 1L))
   )
   verbose
+}
+
+#' Validate `keep_data`
+#'
+#' Validate `keep_data` in [el_control()].
+#'
+#' @param keep_data A single logical.
+#' @return A single logical.
+#' @noRd
+validate_keep_data <- function(keep_data) {
+  stopifnot(
+    "`keep_data` must be a single logical." =
+      (isTRUE(is.logical(keep_data) && length(keep_data) == 1L))
+  )
+  keep_data
 }
 
 #' Validate `nthreads`
@@ -211,23 +228,6 @@ validate_m <- function(m) {
     "`m` must be a positive single integer." = (m > 0L)
   )
   m
-}
-
-#' Validate `model`
-#'
-#' Validate `model` in [el_glm()], [el_lm()], and [el_mean()].
-#'
-#' @param model A single logical.
-#' @return A single logical.
-#' @srrstats {G3.0} The package does not compare floating point numbers for
-#'   equality. All numeric equality comparisons are made between integers.
-#' @noRd
-validate_model <- function(model) {
-  stopifnot(
-    "`model` must be a single logical." =
-      (isTRUE(is.logical(model) && length(model) == 1L))
-  )
-  model
 }
 
 #' Validate `x`

### 2022-07-08 R/confreg-methods.R
methods for `el_sd()` added.

@@ -9,41 +9,40 @@ setMethod("confreg", "EL", function(object,
                                     control = el_control()) {
   level <- validate_level(level)
   npoints <- validate_npoints(npoints)
-  if (!is(control, "ControlEL")) {
-    stop("invalid `control` specified.")
-  }
-  est <- coef(object)
-  if (length(est) == 0L) {
+  stopifnot(
+    "Invalid `control` specified." = (is(control, "ControlEL"))
+  )
+  if (object@npar == 0L) {
     stop("`confreg` method is not applicable to an empty model.")
-  } else if (length(est) == 1L) {
+  } else if (object@npar == 1L) {
     stop("`confreg` method is not applicable to a model with one parameter.")
   }
+  est <- coef(object)
   pnames <- if (is.null(names(est))) seq(length(est)) else names(est)
   if (!missing(parm)) {
-    if (length(parm) != 2L) {
-      stop("length of `parm` must be two.")
-    }
+    stopifnot("Length of `parm` must be two." = (length(parm) == 2L))
     if (is.numeric(parm) && all(is.finite(parm))) {
       idx <- as.integer(parm)
       est <- est[idx]
       pnames <- pnames[idx]
-      if (length(unique(est)) != 2L) {
-        stop("only one parameter specified by `parm`.")
-      }
+      stopifnot(
+        "Specify two parameters for `parm`." =
+          (length(unique(est)) == 2L)
+      )
     } else if (is.character(parm)) {
-      if (is.null(names(est))) {
-        stop(
-          "`parm` is not recognized since 'object' has no named parameters."
-        )
-      }
+      stopifnot(
+        "`parm` is not recognized since 'object' has no named parameters." =
+          (isFALSE(is.null(names(est))))
+      )
       idx <- match(parm, pnames)
       est <- est[idx]
       pnames <- pnames[idx]
-      if (length(unique(est)) != 2L) {
-        stop("only one parameter specified by `parm`.")
-      }
+      stopifnot(
+        "Only one parameter specified by `parm`." =
+          (length(unique(est)) == 2L)
+      )
     } else {
-      stop("invalid `parm` specified.")
+      stop("Invalid `parm` specified.")
     }
   } else {
     idx <- c(1L, 2L)

### 2022-07-08 R/el_sd.R

@@ -5,7 +5,7 @@
 #' @param x A numeric vector, or an object that can be coerced to a numeric
 #'   vector.
 #' @param mean A single numeric for the (known) mean value.
-#' @param sd A single numeric for the parameter value to be tested.
+#' @param sd A positive single numeric for the parameter value to be tested.
 #' @param weights An optional numeric vector of weights to be used in the
 #'   fitting process. The length of the vector must be the same as the length of
 #'   `x`. Defaults to `NULL`, corresponding to identical weights. If non-`NULL`,
@@ -27,17 +27,17 @@ el_sd <- function(x, mean, sd, weights = NULL, control = el_control()) {
   stopifnot(
     "`x` must have at least two observations." = (length(x) >= 2L),
     "`x` must be a finite numeric vector" = (isTRUE(all(is.finite(x)))),
+    "`mean` must be a finite single numeric." =
+      (isTRUE(is.numeric(mean) && length(mean) == 1L && is.finite(mean))),
     "`sd` must be a finite single numeric." =
       (isTRUE(is.numeric(sd) && length(sd) == 1L && is.finite(sd))),
     "`sd` must be a positive single numeric." = (sd >= .Machine$double.eps),
-    "`mean` must be a finite single numeric." =
-      (isTRUE(is.numeric(mean) && length(mean) == 1L && is.finite(mean))),
     "Invalid `control` specified." = (is(control, "ControlEL"))
   )
   n <- length(x)
-  g <- (x - mean)^2L - sd^2
+  mm <- (x - mean)^2L
   if (length(nm) == n) {
-    names(g) <- nm
+    names(mm) <- nm
   }
   w <- validate_weights(weights, n)
   if (length(w) == 0L) {
@@ -46,7 +46,7 @@ el_sd <- function(x, mean, sd, weights = NULL, control = el_control()) {
     est <- sqrt(sum((x - mean)^2L * w) / n)
     names(w) <- nm
   }
-  out <- compute_generic_EL(g, control@maxit_l, control@tol_l, control@th, w)
+  out <- compute_EL("sd", sd, mm, control@maxit_l, control@tol_l, control@th, w)
   optim <- out$optim
   names(optim$sd) <- names(sd)
   if (control@verbose) {
@@ -55,11 +55,11 @@ el_sd <- function(x, mean, sd, weights = NULL, control = el_control()) {
       if (out$optim$convergence) "achieved." else "failed."
     )
   }
-  new("EL",
[…]

### 2022-07-08 src/EL.cpp

@@ -8,7 +8,7 @@
 #include <utility>
 
 /* EL class (evaluation)
- * Last updated: 07/04/22
+ * Last updated: 07/07/22
  */
 EL::EL(const std::string method,
        const Eigen::Ref<const Eigen::VectorXd> &par0,
@@ -56,6 +56,7 @@ EL::set_mele_fn(const std::string method)
                             const Eigen::Ref<const Eigen::MatrixXd> &,
                             const Eigen::Ref<const Eigen::ArrayXd> &)>>
       mele_map{{{"mean", mele_mean},
+                {"sd", mele_sd},
                 {"lm", mele_lm}}};
   return mele_map[method];
 }
@@ -68,6 +69,7 @@ EL::set_g_fn(const std::string method)
                             const Eigen::Ref<const Eigen::MatrixXd> &,
                             const Eigen::Ref<const Eigen::VectorXd> &)>>
       g_map{{{"mean", g_mean},
+             {"sd", g_sd},
              {"lm", g_lm},
              {"gaussian_identity", g_lm},
              {"gaussian_log", g_gauss_log},
@@ -260,6 +262,7 @@ CEL::set_g_fn(const std::string method)
                             const Eigen::Ref<const Eigen::MatrixXd> &,
                             const Eigen::Ref<const Eigen::VectorXd> &)>>
       g_map{{{"mean", g_mean},
+             {"sd", g_sd},
              {"lm", g_lm},
              {"gaussian_identity", g_lm},
              {"gaussian_log", g_gauss_log},
@@ -290,6 +293,7 @@ CEL::set_gr_fn(const std::string method)
                             const bool)>>
       gr_map{
           {{"mean", gr_nloglr_mean},
+           {"sd", gr_nloglr_sd},
            {"lm", gr_nloglr_lm},
            {"gaussian_identity", gr_nloglr_lm},
            {"gaussian_log", gr_nloglr_gauss_log},

### 2022-07-08 src/utils.h

@@ -10,6 +10,8 @@ g_fn2(const std::string method);
 
 Eigen::VectorXd mele_mean(const Eigen::Ref<const Eigen::MatrixXd> &x,
                           const Eigen::Ref<const Eigen::ArrayXd> &w);
+Eigen::VectorXd mele_sd(const Eigen::Ref<const Eigen::MatrixXd> &x,
+                        const Eigen::Ref<const Eigen::ArrayXd> &w);
 Eigen::VectorXd mele_lm(const Eigen::Ref<const Eigen::MatrixXd> &x,
                         const Eigen::Ref<const Eigen::ArrayXd> &w);
 
@@ -30,6 +32,16 @@ Eigen::VectorXd gr_nloglr_mean(const Eigen::Ref<const Eigen::VectorXd> &l,
                                const Eigen::Ref<const Eigen::VectorXd> &par,
                                const Eigen::Ref<const Eigen::ArrayXd> &w,
                                const bool weighted);
+
+Eigen::MatrixXd g_sd(const Eigen::Ref<const Eigen::MatrixXd> &x,
+                     const Eigen::Ref<const Eigen::VectorXd> &par);
+
+Eigen::VectorXd gr_nloglr_sd(const Eigen::Ref<const Eigen::VectorXd> &l,
+                             const Eigen::Ref<const Eigen::MatrixXd> &g,
+                             const Eigen::Ref<const Eigen::MatrixXd> &x,
+                             const Eigen::Ref<const Eigen::VectorXd> &par,
+                             const Eigen::Ref<const Eigen::ArrayXd> &w,
+                             const bool weighted);
 
 Eigen::MatrixXd g_lm(const Eigen::Ref<const Eigen::MatrixXd> &x,
                      const Eigen::Ref<const Eigen::VectorXd> &par);

### 2022-07-09 README.Rmd
adding vignette
README updated

@@
 A collection of functions are available for regression analysis and hypothesis testing.
-Much of its functionality and syntax mimics the corresponding base R functions.
 The core computational routines are implemented with the 'Eigen' C++ library and 'RcppEigen' interface, with OpenMP for parallel computation.
-Additional functions are available for multiple testing for the analysis of experimental designs.
 Details of the testing procedures are given in [Kim, MacEachern, and Peruggia (2021)](https://arxiv.org/abs/2112.09206).
@@
 ## Usage
-```{r eval = TRUE}
+```{r usage, eval = TRUE}
 library(melt)
-# one sample test for mean
-fit1 <- el_mean(rnorm(n = 100), par = 0)
-confint(fit1)
 
-# linear regression
-fit2 <- el_lm(mpg ~ wt, data = mtcars)
+# Test for the mean
+data("precip")
+fit <- el_mean(precip, par = 30)
+fit
+
+# Linear regression
+data("mtcars")
+fit2 <- el_lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
 summary(fit2)
-confint(fit2)
+cr <- confreg(fit2, parm = c("disp", "hp"), npoints = 200)
+plot(cr)
 
-# analysis of variance 
+# Analysis of variance 
 data("clothianidin") 
@@
 
-# test of no treatment effect
+# Test of no treatment effect
 lhs <- matrix(c(1, -1, 0, 0,
@@
 elt(fit3, lhs = lhs)
-
-# logistic regression
-fit4 <- el_glm(vs ~ mpg, data = mtcars, family = binomial)
-summary(fit4)
 ```

### 2022-07-11 README.Rmd
README updated

@@
 A collection of functions are available for regression analysis and hypothesis testing.
-The core computational routines are implemented with the 'Eigen' C++ library and 'RcppEigen' interface, with OpenMP for parallel computation.
+The core computational routines are implemented with the [Eigen](https://eigen.tuxfamily.org/index.php?title=Main_Page) C++ library and [RcppEigen](https://cran.r-project.org/web/packages/RcppEigen/index.html) interface, with OpenMP for parallel computation.
 Details of the testing procedures are given in [Kim, MacEachern, and Peruggia (2021)](https://arxiv.org/abs/2112.09206).

### 2022-07-31 R/print-methods.R
`chisq()` added

@@ -17,10 +17,10 @@ setMethod("print", "EL", function(x,
   cat("\n")
 
   out <- character()
-  if (length(x@statistic) != 0L) {
+  if (length(chisq(x)) != 0L) {
     out <- c(
-      out, paste("Chisq:", format.default(x@statistic, digits = digits)),
-      paste("df:", x@df),
+      out, paste("Chisq:", format.default(chisq(x), digits = digits)),
+      paste("df:", getDF(x)),
       paste("Pr(>Chisq):", format.pval(pVal(x), digits = digits))
     )
   } else {
@@ -28,7 +28,7 @@ setMethod("print", "EL", function(x,
   }
   cat(strwrap(paste(out, collapse = ", ")), sep = "\n\n")
 
-  if (length(x@statistic) != 0L) {
+  if (length(chisq(x)) != 0L) {
     cat(
       "\nEL evaluation:",
       if (conv(x)) "converged" else "not converged", "\n"
@@ -56,10 +56,10 @@ setMethod("print", "LM", function(x,
   cat("\n")
 
   out <- character()
-  if (length(x@statistic) != 0L) {
+  if (length(chisq(x)) != 0L) {
     out <- c(
-      out, paste("Chisq:", format.default(x@statistic, digits = digits)),
-      paste("df:", x@df),
+      out, paste("Chisq:", format.default(chisq(x), digits = digits)),
+      paste("df:", getDF(x)),
       paste("Pr(>Chisq):", format.pval(pVal(x), digits = digits))
     )
   } else {
@@ -67,7 +67,7 @@ setMethod("print", "LM", function(x,
   }
   cat(strwrap(paste(out, collapse = ", ")), sep = "\n\n")
 
-  if (length(x@statistic) != 0L) {
+  if (length(chisq(x)) != 0L) {
     if (x@misc$intercept) {
       cat(
         "\nConstrained EL:",
@@ -106,12 +106,12 @@ setMethod(
       cat("\nNo Coefficients\n")
     } else {
       cat("\nCoefficients:\n")
-      coefs <- getSigTests(x)
+      coefs <- sigTests(x)
       if (any(aliased <- x@aliased)) {
         cn <- names(aliased)
         coefs <-
           matrix(NA, length(aliased), 3L, dimnames = list(cn, colnames(coefs)))
-        coefs[!aliased, ] <- getSigTests(x)
+        coefs[!aliased, ] <- sigTests(x)
       }
       printCoefmat(coefs,
[…]

### 2022-07-31 R/sigTests-methods.R

@@ -1,0 +1,12 @@
+#' @describeIn sigTests Extracts a list with the optimization results of
+#'   significance tests.
+setMethod("sigTests", "LM", function(object, ...) {
+  object@sigTests
+})
+
+
+#' @describeIn sigTests Extracts a matrix with the results of significance
+#'   tests.
+setMethod("sigTests", "SummaryLM", function(object, ...) {
+  object@sigTests
+})

### 2022-08-02 README.Rmd
funding information added
setting up tinytex in github action

--- 2022-07-09 README.Rmd
@@
 A collection of functions are available for regression analysis and hypothesis testing.
-The core computational routines are implemented with the 'Eigen' C++ library and 'RcppEigen' interface, with OpenMP for parallel computation.
+The core computational routines are implemented with the [Eigen](https://eigen.tuxfamily.org/index.php?title=Main_Page) C++ library and [RcppEigen](https://cran.r-project.org/package=RcppEigen) interface, with OpenMP for parallel computation.
 Details of the testing procedures are given in [Kim, MacEachern, and Peruggia (2021)](https://arxiv.org/abs/2112.09206).
+This work was supported by the U.S.
+National Science Foundation under Grants No.
+[SES-1921523](https://nsf.gov/awardsearch/showAward?AWD_ID=1921523) and [DMS-2015552](https://nsf.gov/awardsearch/showAward?AWD_ID=2015552).
 

### 2022-08-24 README.Rmd
bump to 1.7.0.9000
pdf vignette updated

@@
   comment = "#>",
+  dpi = 300,
   fig.path = "man/figures/README-",
@@
 # Test of no treatment effect
-lhs <- matrix(c(1, -1, 0, 0,
+contrast <- matrix(c(
+  1, -1, 0, 0,
   0, 1, -1, 0,
-                0, 0, 1, -1), byrow = TRUE, nrow = 3)
-elt(fit3, lhs = lhs)
+  0, 0, 1, -1
+), byrow = TRUE, nrow = 3)
+elt(fit3, lhs = contrast)
 ```

### 2022-08-24 R/calibrate.R

@@ -7,7 +7,7 @@
 #' @param calibrate A single character.
 #' @param p A single integer.
 #' @param par A numeric vector.
-#' @param object An object that inherit from class \linkS4class{EL}.
+#' @param object An object thats inherit from class \linkS4class{EL}.
 #' @param control An object of class of \linkS4class{ControlEL}.
 #' @return A numeric vector of length two for the critical value and the
 #'   p-value.

### 2022-08-24 R/data-thiamethoxam.R

@@ -1,0 +1,28 @@
+#' Thiamethoxam applications in squash crops
+#'
+#' A dataset on the effect of the thiamethoxam application method and plant
+#'   variety on bees.
+#'
+#' @format A data frame with 55 observations and 11 variables:
+#' \describe{
+#' \item{trt}{Treatment.}
+#' \item{var}{Variety.}
+#' \item{rep}{Replicate.}
+#' \item{fruit}{Average fruit number per plant.}
+#' \item{avg_mass}{Individual Fruit mass average (g).}
+#' \item{mass}{Fruit mass per plant (g).}
+#' \item{yield}{Yield (4 plants).}
+#' \item{visit}{Bee visits per plot.}
+#' \item{foliage}{Proportion of foliage consumed by striped cucumber beetle.}
+#' \item{scb}{Striped cucumber beetle per plant.}
+#' \item{defoliation}{Defoliation percentage.}
+#' }
+#' @usage data("thiamethoxam")
+#' @examples
+#' data("thiamethoxam")
+#' thiamethoxam
+#' @source Obregon D, Pederson G, Taylor A, Poveda K (2022). “The Pest Control
+#'   and Pollinator Protection Dilemma: The Case of Thiamethoxam Prophylactic
+#'   Applications in Squash Crops.” \emph{PLOS ONE}, 17(5), 1–18.
+#'   \doi{10.1371/journal.pone.0267984}.
+"thiamethoxam"

### 2022-08-24 R/sigTests-methods.R

@@ -4,7 +4,6 @@ setMethod("sigTests", "LM", function(object, ...) {
   object@sigTests
 })
 
-
 #' @describeIn sigTests Extracts a matrix with the results of significance
 #'   tests.
 setMethod("sigTests", "SummaryLM", function(object, ...) {

### 2022-08-24 src/EL.cpp

@@ -156,8 +156,9 @@ double EL::loglik() const
   }
 }
 
+
 /* CEL class (minimization)
- * Last updated: 07/04/22
+ * Last updated: 08/23/22
  */
 CEL::CEL(const std::string method,
          const Eigen::Ref<const Eigen::VectorXd> &par0,
@@ -178,12 +179,9 @@ CEL::CEL(const std::string method,
       gr_fn{CEL::set_gr_fn(method)}
 {
   /// initialization ///
-  // orthogonal projection matrix
-  const Eigen::MatrixXd proj =
-      Eigen::MatrixXd::Identity(lhs.cols(), lhs.cols()) -
-      lhs.transpose() * (lhs * lhs.transpose()).inverse() * lhs;
   // parameter (constraint imposed)
-  par = proj * par0 + lhs.transpose() * (lhs * lhs.transpose()).inverse() * rhs;
+  par = proj(lhs, par0) +
+        lhs.bdcSvd(Eigen::ComputeThinU | Eigen::ComputeThinV).solve(rhs);
   // estimating function
   Eigen::MatrixXd g = g_fn(x, par);
   // lambda
@@ -191,14 +189,14 @@ CEL::CEL(const std::string method,
   // function value (-logLR)
   nllr = PseudoLog::sum(Eigen::VectorXd::Ones(n) + g * l, wt);
   // function norm
-  // const double norm0 = (proj * gr_fn(l, g, x, par, wt, weighted)).norm();
+  // const double norm0 = proj(lhs, gr_fn(l, g, x, par, wt, weighted)).norm();
 
   /// minimization (projected gradient descent) ///
   while (!conv && iter != maxit && nllr <= th)
   {
     // update parameter
     Eigen::VectorXd par_tmp =
-        par - gamma * proj * gr_fn(l, g, x, par, wt, weighted);
+        par - gamma * proj(lhs, gr_fn(l, g, x, par, wt, weighted));
     // update estimating function
     Eigen::MatrixXd g_tmp = g_fn(x, par_tmp);
     // update lambda
@@ -214,11 +212,10 @@ CEL::CEL(const std::string method,
       gamma /= 2;
       if (gamma < DBL_EPSILON)
       {
-        // Rcpp::Rcout << "seffff" << "\n";
         break;
       }
       // propose new parameter
[…]

### 2022-08-24 src/EL.h

@@ -126,6 +126,15 @@ private:
       gr_fn;
 };
 
+/* PseudoLog class
+ * The constructor takes an input vector `x` and an optional weight vector `w`.
+ * Three member variables are computed: `dplog`, `sqrt_neg_d2plog`, and
+ * `plog_sum`.
+ * 1. `dplog`: first derivatives of the pseudo-log values.
+ * 2. `sqrt_neg_d2plog`: square root of negative second derivatives of the
+ * pseudo-log values.
+ * 3. `plog_sum`: sum of the pseudo-log values.
+ */
 class PseudoLog
 {
 public:

### 2022-08-24 src/deprecated.cpp

@@ -35,8 +35,8 @@ Rcpp::List pairwise(const Eigen::MatrixXd &x,
     lhs(pairs[i][0]) = 1;
     lhs(pairs[i][1]) = -1;
     minEL pairwise_result =
-        test_gbd_EL(theta_hat, x, c, lhs, Eigen::Matrix<double, 1, 1>(0), th,
-                    maxit, abstol);
+        test_gbd_EL(theta_hat, x, c, lhs, Eigen::Matrix<double, 1, 1>(0),
+                    th*100, maxit, abstol);
     statistic[i] = 2 * pairwise_result.nllr;
     convergence[i] = pairwise_result.convergence;
   }
@@ -52,7 +52,7 @@ Rcpp::List pairwise(const Eigen::MatrixXd &x,
   else
   {
     bootstrap_statistics_pairwise = bootstrap_statistics_pairwise_NB(
-        x, c, k, pairs, B, level, nthreads, th, maxit, abstol);
+        x, c, k, pairs, B, level, nthreads, th*100, maxit, abstol);
   }
   std::vector<double> adj_pvalues(m);
   for (int i = 0; i < m; ++i)
@@ -87,7 +87,7 @@ Rcpp::List pairwise(const Eigen::MatrixXd &x,
       lhs(pairs[i][0]) = 1;
       lhs(pairs[i][1]) = -1;
       std::array<double, 2> ci = pair_confidence_interval_gbd(
-          theta_hat, x, c, lhs, th, estimate[i], cutoff);
+          theta_hat, x, c, lhs, th*100, estimate[i], cutoff);
       lower[i] = ci[0];
       upper[i] = ci[1];
     }

### 2022-08-24 src/helpers.cpp

@@ -28,6 +28,12 @@ set_g_fn(const std::string method)
   return g_map[method];
 }
 
+Eigen::VectorXd proj(const Eigen::Ref<const Eigen::MatrixXd> &l,
+                     const Eigen::Ref<const Eigen::VectorXd> &x)
+{
+  return x - l.transpose() * ((l * l.transpose()).householderQr().solve(l * x));
+}
+
 Eigen::ArrayXd inverse_linkinv(const Eigen::Ref<const Eigen::VectorXd> &x)
 {
   return x.array().inverse();

### 2022-08-25 README.Rmd
README updated

@@
 ## Overview
-The R package **melt** provides a unified framework for data analysis with empirical likelihood methods.
-A collection of functions are available for regression analysis and hypothesis testing.
+Empirical likelihood enables a nonparametric, likelihood-driven style of inference without relying on assumptions frequently made in parametric models.
+Empirical likelihood-based tests are asymptotically pivotal and thus avoid explicit studentization, making it also challenging for the methods to be directly extended to other packages for parametric models.
+The R package [melt](https://cran.r-project.org/package=melt) aims to bridge the gap and provide a unified framework for data analysis with empirical likelihood methods.
+A collection of functions are available to perform multiple empirical likelihood tests and construct confidence intervals for linear and generalized linear models in R.
+The package offers an easy-to-use interface and flexibility in specifying hypotheses and calibration methods, extending the framework to simultaneous inference.
 The core computational routines are implemented with the [Eigen](https://eigen.tuxfamily.org/index.php?title=Main_Page) C++ library and [RcppEigen](https://cran.r-project.org/package=RcppEigen) interface, with OpenMP for parallel computation.

### 2022-08-25 R/AllClasses.R

@@ -3,8 +3,8 @@
 #' S4 class for empirical likelihood.
 #'
 #' @details Let \eqn{X_i} be independent and identically distributed
-#'   \eqn{p}-dimensional random variable from an unknown distribution \eqn{F}
-#'   for \eqn{i = 1, \dots, n}. We assume that \eqn{F} has a positive definite
+#'   \eqn{p}-dimensional random variable from an unknown distribution \eqn{P}
+#'   for \eqn{i = 1, \dots, n}. We assume that \eqn{P} has a positive definite
 #'   covariance matrix. For a parameter of interest
 #'   \eqn{\theta(F) \in {\rm{I\!R}}^p}, consider a \eqn{p}-dimensional smooth
 #'   estimating function \eqn{g(X_i, \theta)} with a moment condition
@@ -12,7 +12,7 @@
 #'   We assume that there exists an unique \eqn{\theta_0} that solves the above
 #'   equation. Given a value of \eqn{\theta}, the (profile) empirical likelihood
 #'   ratio is defined by
-#'   \deqn{\mathcal{R}(\theta) =
+#'   \deqn{R(\theta) =
 #'   \max_{p_i}\left\{\prod_{i = 1}^n np_i :
 #'   \sum_{i = 1}^n p_i g(X_i, \theta) = 0, p_i \geq 0, \sum_{i = 1}^n p_i = 1
 #'   \right\}.}
@@ -23,16 +23,16 @@
 #'   \deqn{\frac{1}{n}\sum_{i = 1}^n \frac{g(X_i, \theta)}
 #'   {1 + \lambda^\top g(X_i, \theta)} = 0.}
 #'   Then the empirical log-likelihood ratio is given by
-#'   \deqn{\log\mathcal{R}(\theta) = -\sum_{i = 1}^n
+#'   \deqn{\log R(\theta) = -\sum_{i = 1}^n
 #'   \log(1 + \lambda^\top g(X_i, \theta)).}
 #'   This problem can be efficiently solved by the Newton-Raphson method when
 #'   the zero vector is contained in the interior of the convex hull of
 #'   \eqn{\{g(X_i, \theta)\}_{i = 1}^n}.
 #'
-#'   Under some regularity conditions, it is known that
-#'   \eqn{-2\log\mathcal{R}(\theta_0)} converges in distribution to
-#'   \eqn{\chi^2_p}, where \eqn{\chi^2_p} has a chi-square distribution with
-#'   \eqn{p} degrees of freedom.
[…]

### 2022-08-25 R/AllGenerics.R

@@ -59,7 +59,7 @@ setGeneric("coef", function(object, ...) standardGeneric("coef"))
 #'  respectively.
 #' @references Owen A (1990).
 #'   “Empirical Likelihood Ratio Confidence Regions.”
-#'   \emph{The Annals of Statistics}, 18(1), 90–120.
+#'   \emph{The Annals of Statistics}, 18(1), 90--120.
 #'   \doi{10.1214/aos/1176347494}.
 #' @seealso [confreg()], [el_control()], [elt()]
 #' @usage NULL
@@ -97,7 +97,7 @@ setGeneric("confint", function(object, parm, level = 0.95, ...)
 #' @return An object of class \linkS4class{ConfregEL}.
 #' @references Owen A (1990).
 #'   “Empirical Likelihood Ratio Confidence Regions.”
-#'   \emph{The Annals of Statistics}, 18(1), 90–120.
+#'   \emph{The Annals of Statistics}, 18(1), 90--120.
 #'   \doi{10.1214/aos/1176347494}.
 #' @seealso [confint()], [el_control()], [elt()], [plot()]
 #' @usage NULL
@@ -182,7 +182,7 @@ setGeneric("critVal", function(object, ...) standardGeneric("critVal"))
 #'   \doi{10.1198/106186005X59568}.
 #' @references Zhu H, Ibrahim JG, Tang N, Zhang H (2008).
 #'   “Diagnostic Measures for Empirical Likelihood of General Estimating
-#'   Equations.” \emph{Biometrika}, 95(2), 489–507.
+#'   Equations.” \emph{Biometrika}, 95(2), 489--507.
 #'   \doi{10.1093/biomet/asm094}.
 #' @seealso [el_control()], [el_eval()], [plot()]
 #' @usage NULL
@@ -317,11 +317,11 @@ setGeneric("elmt", function(object,
 #'   corresponds to that of \linkS4class{EL}.
 #' @references Adimari G, Guolo A (2010).
 #'   “A Note on the Asymptotic Behaviour of Empirical Likelihood Statistics.”
-#'   \emph{Statistical Methods & Applications}, 19(4), 463–476.
+#'   \emph{Statistical Methods & Applications}, 19(4), 463--476.
 #'   \doi{10.1007/s10260-010-0137-9}.
 #' @references Qin J, Lawless J (1995).
 #'   “Estimating Equations, Empirical Likelihood and Constraints on Parameters.”
[…]

### 2022-08-25 R/data-clothianidin.R

@@ -16,8 +16,7 @@
 #' @usage data("clothianidin")
 #' @examples
 #' data("clothianidin")
-#' clothianidin
 #' @source Alford A, Krupke CH (2017). “Translocation of the Neonicotinoid Seed
-#'   Treatment Clothianidin in Maize.” \emph{PLOS ONE}, 12(3), 1–19.
+#'   Treatment Clothianidin in Maize.” \emph{PLOS ONE}, 12(3), 1--19.
 #'   \doi{10.1371/journal.pone.0173836}.
 "clothianidin"

### 2022-08-25 R/el_eval.R

@@ -11,10 +11,10 @@
 #'   non-`NULL`, weighted empirical likelihood is computed.
 #' @param control An object of class \linkS4class{ControlEL} constructed by
 #'   [el_control()].
-#' @details `el_eval` evaluates empirical likelihood with a \eqn{n} by
+#' @details [el_eval()] evaluates empirical likelihood with a \eqn{n} by
 #'   \eqn{p} numeric matrix argument `g`, whose \eqn{i}th row is
 #'   \eqn{g(X_i, \theta)}. Since the estimating function can be arbitrary,
-#'   `el_eval` does not return an object of class \linkS4class{EL}, and the
+#'   [el_eval()] does not return an object of class \linkS4class{EL}, and the
 #'   associated generics and methods are not applicable.
 #' @return A list with the following components:
 #'   * `optim` A list with the following optimization results:
@@ -32,9 +32,9 @@
 #'   * `weights` Re-scaled weights used for model fitting.
 #' @references Qin J, Lawless J (1994).
 #'   “Empirical Likelihood and General Estimating Equations.”
-#'   \emph{The Annals of Statistics}, 22(1), 300–325.
+#'   \emph{The Annals of Statistics}, 22(1), 300--325.
 #'   \doi{10.1214/aos/1176325370}.
-#' @seealso [el_control()]
+#' @seealso \linkS4class{EL}, [el_control()]
 #' @examples
 #' set.seed(3271)
 #' x <- rnorm(50)

### 2022-09-08 README.Rmd
documentation updated

@@
 
+## Main functions
+melt provides an intuitive API for performing the most common data analysis tasks:
+
+* `el_lm()` fits a linear model with empirical likelihood.
+* `el_glm()` fits a generalized linear model with empirical likelihood.
+* `confint()` computes confidence intervals for model parameters.
+* `confreg()` computes a confidence region for model parameters.
+* `elt()` tests a linear hypothesis.
+* `elmt()` performs multiple testing simultaneously.
+
 ## Usage
@@
 library(melt)
+set.seed(971112)
 
-# Test for the mean
+## Test for the mean
 data("precip")
-fit <- el_mean(precip, par = 30)
-fit
+el_mean(precip, par = 30)
 
-# Linear regression
+## Linear model
 data("mtcars")
-fit2 <- el_lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
-summary(fit2)
-cr <- confreg(fit2, parm = c("disp", "hp"), npoints = 200)
+fit_lm <- el_lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
+summary(fit_lm)
+cr <- confreg(fit_lm, parm = c("disp", "hp"), npoints = 200)
 plot(cr)
+data("clothianidin")
+fit2_lm <- el_lm(clo ~ -1 + trt, data = clothianidin)
+summary(fit2_lm)
+confint(fit2_lm)
 
-# Analysis of variance
-data("clothianidin")
-fit3 <- el_lm(clo ~ -1 + trt, data = clothianidin)
-summary(fit3)
-confint(fit3)
+## Generalized linear model
+data("thiamethoxam")
+fit_glm <- el_glm(visit ~ log(mass) + fruit + foliage + var + trt,
+  family = quasipoisson(link = "log"), data = thiamethoxam,
+  control = el_control(maxit = 100, tol = 1e-08, nthreads = 4)
+)
+summary(fit_glm)
 
-# Test of no treatment effect
-contrast <- matrix(c(
-  1, -1, 0, 0,
-  0, 1, -1, 0,
-  0, 0, 1, -1
-), byrow = TRUE, nrow = 3)
-elt(fit3, lhs = contrast)
+## Test of no treatment effect
+contrast <- c(
+  "trtNaked - trtFungicide", "trtFungicide - trtLow", "trtLow - trtHigh"
+)
+elt(fit2_lm, lhs = contrast)
+
+## Multiple testing
+contrast2 <- rbind(
+  c(0, 0, 0, 0, 0, 1, 0, 0),
+  c(0, 0, 0, 0, 0, 0, 1, 0),
+  c(0, 0, 0, 0, 0, 0, 0, 1)
+)
+elmt(fit_glm, lhs = contrast2)
 ```

### 2022-09-08 tests/testthat/setup.R

@@ -7,5 +7,6 @@ data("mtcars")
 data("precip")
 data("sleep")
 data("trees")
+data("warpbreaks")
 data("women")
 set.seed(10358)

### 2022-09-14 README.Rmd
`elmt()` accepts a chracter vector as an  element of `lhs`

@@
 A collection of functions are available to perform multiple empirical likelihood tests and construct confidence intervals for linear and generalized linear models in R.
-The package offers an easy-to-use interface and flexibility in specifying hypotheses and calibration methods, extending the framework to simultaneous inference.
+The package offers an easy-to-use interface and flexibility in specifying hypotheses and calibration methods, extending the framework to simultaneous inferences.
 The core computational routines are implemented with the [Eigen](https://eigen.tuxfamily.org/index.php?title=Main_Page) C++ library and [RcppEigen](https://cran.r-project.org/package=RcppEigen) interface, with OpenMP for parallel computation.

### 2022-09-14 R/print-methods.R

@@ -67,13 +67,17 @@ setMethod("print", "ELMT", function(x,
     P.values = TRUE, has.Pvalue = TRUE, eps.Pvalue = 1e-03
   )
   cat("\n")
-  cat(paste("Common critical value:", round(critVal(x), digits = 4L)), "\n\n")
+  cat(paste(
+    "Common critical value:", format.default(critVal(x), digits = digits)
+  ), "\n\n")
   invisible(x)
 })
 setMethod("show", "ELMT", function(object) print(object))
 
 #' @rdname print
-setMethod("print", "ELT", function(x, digits = getOption("digits"), ...) {
+setMethod("print", "ELT", function(x,
+                                   digits = max(3L, getOption("digits") - 3L),
+                                   ...) {
   cat("\n\tEmpirical Likelihood Test\n\n")
   method <- switch(x@calibrate,
     "chisq" = "Chi-square",
@@ -81,7 +85,7 @@ setMethod("print", "ELT", function(x, digits = getOption("digits"), ...) {
     "f" = "F"
   )
   cat("Hypothesis:", describe_hypothesis(
-    x@rhs, x@lhs, names(getOptim(x)$par)[seq_len(ncol(x@lhs))]
+    x@rhs, x@lhs, names(getOptim(x)$par)[seq_len(ncol(x@lhs))], digits
   ),
   sep = "\n"
   )

### 2022-09-14 tests/testthat/test-skip_on_cran.R

@@ -1,3 +1,30 @@
+test_that("Bootstrap calibration for `elt()`.", {
+  fit <- el_mean(precip, par = 60)
+  fit2 <- el_glm(gear ~ 1,
+    family = quasipoisson("log"), data = mtcars
+  )
+  fit3 <- el_glm(gear ~ mpg + disp,
+    family = quasipoisson("identity"), data = mtcars, weights = wt
+  )
+  fit4 <- el_sd(women$height, mean = 65, sd = 5)
+  fit5 <- el_lm(fruit ~ trt, data = thiamethoxam)
+  fit6 <- el_glm(fruit ~ trt, family = gaussian("log"), data = thiamethoxam)
+  fit7 <- el_glm(fruit ~ trt, family = gaussian("inverse"), data = thiamethoxam)
+  fit8 <- el_glm(gear ~ mpg + disp, family = poisson("sqrt"), data = mtcars)
+  fit9 <- el_glm(wool ~ ., family = binomial("logit"), data = warpbreaks)
+  fit10 <- el_glm(wool ~ ., family = binomial("probit"), data = warpbreaks)
+  expect_s4_class(elt(fit, rhs = 65, calibrate = "boot"), "ELT")
+  expect_s4_class(elt(fit2, rhs = coef(fit2), calibrate = "boot"), "ELT")
+  expect_s4_class(elt(fit3, rhs = coef(fit3), calibrate = "boot"), "ELT")
+  expect_s4_class(elt(fit4, rhs = coef(fit4), calibrate = "boot"), "ELT")
+  expect_s4_class(elt(fit5, rhs = coef(fit5), calibrate = "boot"), "ELT")
+  expect_s4_class(elt(fit6, rhs = coef(fit6), calibrate = "boot"), "ELT")
+  expect_s4_class(elt(fit7, rhs = coef(fit7), calibrate = "boot"), "ELT")
+  expect_s4_class(elt(fit8, rhs = coef(fit8), calibrate = "boot"), "ELT")
+  expect_s4_class(elt(fit9, rhs = coef(fit9), calibrate = "boot"), "ELT")
+  expect_s4_class(elt(fit10, rhs = coef(fit10), calibrate = "boot"), "ELT")
+})
+
 test_that("Parallel computation yields the same results (gaussian - log).", {
   skip_on_cran()
   set.seed(23)

### 2022-09-17 README.Rmd
vignette updated

@@
 Empirical likelihood enables a nonparametric, likelihood-driven style of inference without relying on assumptions frequently made in parametric models.
-Empirical likelihood-based tests are asymptotically pivotal and thus avoid explicit studentization, making it also challenging for the methods to be directly extended to other packages for parametric models.
+Empirical likelihood-based tests are asymptotically pivotal and thus avoid explicit studentization.
+For this reason it is challenging to incorporate empirical likelihood methods directly into other packages that perform inferences for parametric models.
 The R package [melt](https://cran.r-project.org/package=melt) aims to bridge the gap and provide a unified framework for data analysis with empirical likelihood methods.

### 2022-09-19 README.Rmd
README updated

@@
 
-# melt - Multiple Empirical Likelihood Tests
+# melt - Multiple Empirical Likelihood Tests <a href="https://markean.github.io/melt/"><img src="man/figures/logo.png" align="right" height="139" /></a>
+
 <!-- badges: start -->
@@
 * `confint()` computes confidence intervals for model parameters.
-* `confreg()` computes a confidence region for model parameters.
+* `confreg()` computes confidence region for model parameters.
 * `elt()` tests a linear hypothesis.

### 2022-09-20 README.Rmd
begin transfer to ropensci

@@
 [![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
+[![R-CMD-check](https://github.com/ropensci/melt/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/ropensci/melt/actions/workflows/check-standard.yaml)
+[![Codecov test coverage](https://codecov.io/gh/markean/melt/branch/master/graph/badge.svg)](https://app.codecov.io/gh/markean/melt?branch=master)
 [![CRAN status](https://www.r-pkg.org/badges/version/melt)](https://CRAN.R-project.org/package=melt)
-[![R-CMD-check](https://github.com/markean/melt/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/markean/melt/actions/workflows/check-standard.yaml)
-[![Codecov test coverage](https://codecov.io/gh/markean/melt/branch/master/graph/badge.svg)](https://app.codecov.io/gh/markean/melt?branch=master)
+[![ropensci review](https://badges.ropensci.org/550_status.svg)](https://github.com/ropensci/software-review/issues/550)
 <!-- badges: end -->
@@
 You can install the latest development version from
-[Github](https://github.com/markean/melt).
+[Github](https://github.com/ropensci/melt).
 ```{r eval = FALSE}
@@
 ```
+
+---
+Please note that this package is released with a [Contributor Code of Conduct](https://ropensci.org/code-of-conduct/). By contributing to this project, you agree to abide by its terms.

### 2022-09-20 README.Rmd
transfer to ropensci

@@
 
-# melt - Multiple Empirical Likelihood Tests <a href="https://markean.github.io/melt/"><img src="man/figures/logo.png" align="right" height="139" /></a>
+# melt - Multiple Empirical Likelihood Tests <a href="https://docs.ropensci.org/melt/"><img src="man/figures/logo.png" align="right" height="139" /></a>
 
@@
 [![R-CMD-check](https://github.com/ropensci/melt/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/ropensci/melt/actions/workflows/check-standard.yaml)
-[![Codecov test coverage](https://codecov.io/gh/markean/melt/branch/master/graph/badge.svg)](https://app.codecov.io/gh/markean/melt?branch=master)
+[![Codecov test coverage](https://codecov.io/gh/ropensci/melt/branch/master/graph/badge.svg)](https://app.codecov.io/gh/ropensci/melt?branch=master)
 [![CRAN status](https://www.r-pkg.org/badges/version/melt)](https://CRAN.R-project.org/package=melt)
+[![runiverse](https://ropensci.r-universe.dev/badges/melt)](https://ropensci.r-universe.dev/ui#package:melt)
 [![ropensci review](https://badges.ropensci.org/550_status.svg)](https://github.com/ropensci/software-review/issues/550)
@@
 You can install the latest development version from
-[Github](https://github.com/ropensci/melt).
+[Github](https://github.com/ropensci/melt) or [R-universe](https://ropensci.r-universe.dev/ui#package:melt).
 ```{r eval = FALSE}
 # install.packages("devtools")
-devtools::install_github("markean/melt")
+devtools::install_github("ropensci/melt")
+
+install.packages("melt", repos = "https://ropensci.r-universe.dev")
 ```

### 2022-09-20 README.Rmd
srr tags removed

@@
 You can install the latest development version from
-[Github](https://github.com/ropensci/melt) or [R-universe](https://ropensci.r-universe.dev/ui#package:melt).
+[GitHub](https://github.com/ropensci/melt) or [R-universe](https://ropensci.r-universe.dev/ui#package:melt).
 ```{r eval = FALSE}
@@
 devtools::install_github("ropensci/melt")
+```
 
+```{r eval = FALSE}
 install.packages("melt", repos = "https://ropensci.r-universe.dev")

### 2022-09-20 R/coef-methods.R

@@ -1,6 +1,5 @@
 #' @describeIn coef Extracts the numeric vector of the maximum empirical
 #'   likelihood estimates.
-#' @srrstats {RE4.2} `coef()` method is exported.
 setMethod("coef", "EL", function(object, ...) {
   object@coefficients
 })

### 2022-09-20 R/confint-methods.R

@@ -1,5 +1,4 @@
 #' @rdname confint
-#' @srrstats {RE4.3} `confint()` method is exported.
 setMethod("confint", "EL", function(object,
                                     parm,
                                     level = 0.95,

### 2022-09-20 R/confreg-methods.R

@@ -1,6 +1,4 @@
 #' @rdname confreg
-#' @srrstats {G2.4, G2.4c} `as.character()` is used to `pnames`, a character
-#'   vector of length two for the name of parameters.
 setMethod("confreg", "EL", function(object,
                                     parm,
                                     level = 0.95,

### 2022-09-20 R/logL-methods.R

@@ -1,6 +1,4 @@
 #' @rdname logL
-#' @srrstats {RE4.11} `logL()` method extracts the empirical log-likelihood
-#'   value which can be interpreted as a measure of goodness-of-fit.
 setMethod("logL", "EL", function(object, ...) {
   object@logl
 })

### 2022-09-20 R/nobs-methods.R

@@ -1,6 +1,4 @@
 #' @rdname nobs
-#' @srrstats {RE4.5} `nobs()` method extracts the number of observations from a
-#'   model.
 setMethod("nobs", "EL", function(object, ...) {
   object@nobs
 })

### 2022-09-20 tests/testthat/setup.R

@@ -1,4 +1,3 @@
-#' @srrstats {G5.0} Tests use data sets provided by the `datasets` package.
 data("airquality")
 data("attenu")
 data("clothianidin")

### 2022-09-20 tests/testthat/test-el_eval.R

@@ -40,7 +40,6 @@ test_that("Conversion between `logl` and `loglr`.", {
   )
 })
 
-#' @srrstats {G5.3} `el_eval()` returns `logp` that has no missing values.
 test_that("No missing values are returned.", {
   x <- women$height
   n <- length(x)

### 2022-09-20 tests/testthat/test-skip_on_cran.R

@@ -229,28 +229,6 @@ test_that("Parallel computation yields the same results (poisson - sqrt).", {
   expect_equal(sigTests(wfit), sigTests(wfit2))
 })
 
-#' @srrstats {G5.4, G5.4a, G5.5} We perform correctness tests with two different
-#'   data sets; one is the `women` data set from the `datasets` package and the
-#'   other is simulated. `set.seed()` is used for the simulated data.
-#'   Specifically, we consider the correctness of `elt()`, one of the main
-#'   functions in the package. This function performs a constrained optimization
-#'   of empirical likelihood. Since this is the first implementation of using
-#'   projected gradient descent to the nested optimization of empirical
-#'   likelihood, we consider a very simple problem with only two parameters.
-#'   The hypothesis states that the second parameter in the model is some fixed
-#'   number (150 for the `women` and 0 for the simulated data set). Then the
-#'   optimization problem reduces to finding an optimal value for the first
-#'   parameter, and this can be done using `optim()` function in the `stats`
-#'   package. We test whether the `eld()` and the `optim()` produces the same
-#'   results in terms of the solution and the convergence status. The tolerance
-#'   for comparing floating point numbers is set to `1e-07`, which we used
-#'   throughout all unit tests in the package. The threshold value `th` for
-#'   empirical likelihood calculation is set to a large number `1e+10`. This
-#'   prevents the optimization routine from exiting early when large empirical
-#'   likelihood value is encountered. Note that it serves as a safeguard and
-#'   does not affect the correctness of the test. No further manual
-#'   specification of the control parameters are made in `el_control()` or
-#'   `optim()`.
 test_that("Correctness tests.", {
   skip_on_cran()
   fit <- el_mean(women, par = colMeans(women))
@@ -276,10 +254,6 @@ test_that("Correctness tests.", {
   expect_equal(getOptim(out3)$par[1], out4$par, tolerance = 1e-07)
 })
 
-#' @srrstats {RE7.1, RE7.1a} We first simulate a design matrix with fixed
-#'   coefficients. Then two response vectors `y1` and `y2` are generated, where
-#'   `y1` has no noise while some noise is added to `y2`. Fitting a model using
-#'   `el_lm()` is faster with `y1` than with `y2`.
[…]

### 2023-05-05 R/deprecated.R
documentation updated

@@ -30,9 +30,10 @@
 #' @param abstol A single numeric for the the absolute convergence tolerance for
 #'   optimization. Defaults to `1e-08`.
 #' @return A list with class \code{c("pairwise", "melt")}.
-#' @references Kim E, MacEachern S, Peruggia M (2021).
+#' @references Kim E, MacEachern SN, Peruggia M (2023).
 #'   “Empirical Likelihood for the Analysis of Experimental Designs.”
-#'   arxiv:2112.09206. URL <https://arxiv.org/abs/2112.09206>.
+#'   \emph{Journal of Nonparametric Statistics}.
+#'   \doi{10.1080/10485252.2023.2206919}.
 #' @examples
 #' \dontrun{
 #' # All pairwise comparisons

### 2023-05-14 README.Rmd
documentation updated
URLs have been fixed

@@
 [![CRAN status](https://www.r-pkg.org/badges/version/melt)](https://CRAN.R-project.org/package=melt)
-[![runiverse](https://ropensci.r-universe.dev/badges/melt)](https://ropensci.r-universe.dev/ui#package:melt)
+[![runiverse](https://ropensci.r-universe.dev/badges/melt)](http://ropensci.r-universe.dev/ui/#package:melt)
 [![ropensci review](https://badges.ropensci.org/550_status.svg)](https://github.com/ropensci/software-review/issues/550)
@@
 The core computational routines are implemented with the [Eigen](https://eigen.tuxfamily.org/index.php?title=Main_Page) C++ library and [RcppEigen](https://cran.r-project.org/package=RcppEigen) interface, with OpenMP for parallel computation.
-Details of the testing procedures are given in [Kim, MacEachern, and Peruggia (2021)](https://arxiv.org/abs/2112.09206). 
+Details of the testing procedures are given in [Kim, MacEachern, and Peruggia (2023)](https://doi.org/10.1080/10485252.2023.2206919). 
 This work was supported by the U.S.
 National Science Foundation under Grants No.
-[SES-1921523](https://nsf.gov/awardsearch/showAward?AWD_ID=1921523) and [DMS-2015552](https://nsf.gov/awardsearch/showAward?AWD_ID=2015552).
+[SES-1921523](https://www.nsf.gov/awardsearch/showAward?AWD_ID=1921523) and [DMS-2015552](https://www.nsf.gov/awardsearch/showAward?AWD_ID=2015552).
 
@@
 You can install the latest development version from
-[GitHub](https://github.com/ropensci/melt) or [R-universe](https://ropensci.r-universe.dev/ui#package:melt).
+[GitHub](https://github.com/ropensci/melt) or [R-universe](http://ropensci.r-universe.dev/ui/#package:melt).
 ```{r eval = FALSE}

### 2024-02-15 README.Rmd
new CITATION added and README updated

@@
 
-# melt - Multiple Empirical Likelihood Tests <a href="https://docs.ropensci.org/melt/"><img src="man/figures/logo.png" align="right" height="139" /></a>
+# melt  
+<a href="https://docs.ropensci.org/melt/"><img src="man/figures/logo.png" align="right" height="139" /></a>
 
@@
 ## Overview
-Empirical likelihood enables a nonparametric, likelihood-driven style of inference without relying on assumptions frequently made in parametric models.
-Empirical likelihood-based tests are asymptotically pivotal and thus avoid explicit studentization.
-For this reason it is challenging to incorporate empirical likelihood methods directly into other packages that perform inferences for parametric models.
-The R package [melt](https://cran.r-project.org/package=melt) aims to bridge the gap and provide a unified framework for data analysis with empirical likelihood methods.
-A collection of functions are available to perform multiple empirical likelihood tests and construct confidence intervals for linear and generalized linear models in R.
-The package offers an easy-to-use interface and flexibility in specifying hypotheses and calibration methods, extending the framework to simultaneous inferences.
-The core computational routines are implemented with the [Eigen](https://eigen.tuxfamily.org/index.php?title=Main_Page) C++ library and [RcppEigen](https://cran.r-project.org/package=RcppEigen) interface, with OpenMP for parallel computation.
+
+melt provides a unified framework for data analysis with empirical likelihood methods.
+A collection of functions are available to perform multiple empirical likelihood tests and construct confidence intervals for various models in 'R'.
+It offers an easy-to-use interface and flexibility in specifying hypotheses and calibration methods, extending the framework to simultaneous inferences.
+The core computational routines are implemented with the 'Eigen' 'C++' library and 'RcppEigen' interface, with 'OpenMP' for parallel computation. 
 Details of the testing procedures are given in [Kim, MacEachern, and Peruggia (2023)](https://doi.org/10.1080/10485252.2023.2206919). 
@@
 ## Installation
-You can install the latest stable release from [CRAN](https://cran.r-project.org/package=melt).
-```{r eval = FALSE}
+
+You can install the latest stable release of melt from CRAN.
+
+```{r, eval = FALSE}
 install.packages("melt")
 ```
-You can install the latest development version from
-[GitHub](https://github.com/ropensci/melt) or [R-universe](http://ropensci.r-universe.dev/ui/#package:melt).
-```{r eval = FALSE}
-# install.packages("devtools")
-devtools::install_github("ropensci/melt")
+
+You can install the development version of melt from GitHub or [R-universe](http://ropensci.r-universe.dev/ui/#package:melt).
+
+```{r, eval = FALSE}
+# install.packages("pak")
+pak::pak("ropensci/melt")
 ```
@@
 ## Main functions
+
 melt provides an intuitive API for performing the most common data analysis tasks:
 
+* `el_mean()` computes empirical likelihood for the mean.
 * `el_lm()` fits a linear model with empirical likelihood.

### 2024-02-15 README.Rmd
new JSS reference added

@@
 melt provides a unified framework for data analysis with empirical likelihood methods.
-A collection of functions are available to perform multiple empirical likelihood tests and construct confidence intervals for various models in 'R'.
-It offers an easy-to-use interface and flexibility in specifying hypotheses and calibration methods, extending the framework to simultaneous inferences.
+A collection of functions is available to perform multiple empirical likelihood tests and construct confidence intervals for various models in 'R'.
+melt offers an easy-to-use interface and flexibility in specifying hypotheses and calibration methods, extending the framework to simultaneous inferences.
 The core computational routines are implemented with the 'Eigen' 'C++' library and 'RcppEigen' interface, with 'OpenMP' for parallel computation. 
-Details of the testing procedures are given in [Kim, MacEachern, and Peruggia (2023)](https://doi.org/10.1080/10485252.2023.2206919). 
+Details of the testing procedures are provided in [Kim, MacEachern, and Peruggia (2023)](https://doi.org/10.1080/10485252.2023.2206919). 
+The package has a companion paper by [Kim, MacEachern, and Peruggia (2024)](https://doi.org/10.18637/jss.v108.i05).
 This work was supported by the U.S.

### 2024-02-15 R/el_eval.R

@@ -49,9 +49,10 @@
 #'   * `npar` A single integer for the number of parameters.
 #'   * `weights` A numeric vector of the re-scaled weights used for the model
 #'   fitting.
-#' @references Qin J, Lawless J (1994).
-#'   “Empirical Likelihood and General Estimating Equations.”
-#'   \emph{The Annals of Statistics}, 22(1), 300--325.
+#' @references
+#'   Qin J, Lawless J (1994).
+#'   ``Empirical Likelihood and General Estimating Equations.''
+#'   \emph{The Annals of Statistics}, **22**(1), 300--325.
 #'   \doi{10.1214/aos/1176325370}.
 #' @seealso \linkS4class{EL}, [el_control()]
 #' @examples

### 2024-02-16 README.Rmd
* `el_pairwise()` and its associated methods are removed.

@@
 
-# melt  
-<a href="https://docs.ropensci.org/melt/"><img src="man/figures/logo.png" align="right" height="139" /></a>
+# melt <a href="https://docs.ropensci.org/melt/"><img src="man/figures/logo.png" align="right" height="139" /></a>
 
@@
 [![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
-[![R-CMD-check](https://github.com/ropensci/melt/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/ropensci/melt/actions/workflows/check-standard.yaml)
+[![R-CMD-check](https://github.com/ropensci/melt/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ropensci/melt/actions/workflows/R-CMD-check.yaml)
+[![pkgcheck](https://github.com/ropensci/melt/workflows/pkgcheck/badge.svg)](https://github.com/ropensci/melt/actions?query=workflow%3Apkgcheck)
 [![Codecov test coverage](https://codecov.io/gh/ropensci/melt/branch/master/graph/badge.svg)](https://app.codecov.io/gh/ropensci/melt?branch=master)
@@
 [![ropensci review](https://badges.ropensci.org/550_status.svg)](https://github.com/ropensci/software-review/issues/550)
+
 <!-- badges: end -->

### 2024-02-16 R/RcppExports.R

@@ -29,10 +29,6 @@ compute_generic_EL <- function(g, maxit_l, tol_l, th, w) {
     .Call(`_melt_compute_generic_EL`, g, maxit_l, tol_l, th, w)
 }
 
-pairwise <- function(x, c, control = 0L, k = 1L, level = 0.05, interval = TRUE, method = "AMC", B = 1e+04L, nthreads = 1L, th = 50, maxit = 1e+04L, abstol = 1e-08) {
-    .Call(`_melt_pairwise`, x, c, control, k, level, interval, method, B, nthreads, th, maxit, abstol)
-}
-
 test_GLM <- function(method, x, par0, intercept, maxit, maxit_l, tol, tol_l, step, th, nthreads, w) {
     .Call(`_melt_test_GLM`, method, x, par0, intercept, maxit, maxit_l, tol, tol_l, step, th, nthreads, w)
 }

### 2024-02-16 src/RcppExports.cpp

@@ -156,28 +156,6 @@ BEGIN_RCPP
     return rcpp_result_gen;
 END_RCPP
 }
-// pairwise
-Rcpp::List pairwise(const Eigen::MatrixXd& x, const Eigen::MatrixXd& c, const int control, const int k, const double level, const bool interval, const std::string method, const int B, const int nthreads, const double th, const int maxit, const double abstol);
-RcppExport SEXP _melt_pairwise(SEXP xSEXP, SEXP cSEXP, SEXP controlSEXP, SEXP kSEXP, SEXP levelSEXP, SEXP intervalSEXP, SEXP methodSEXP, SEXP BSEXP, SEXP nthreadsSEXP, SEXP thSEXP, SEXP maxitSEXP, SEXP abstolSEXP) {
-BEGIN_RCPP
-    Rcpp::RObject rcpp_result_gen;
-    Rcpp::RNGScope rcpp_rngScope_gen;
-    Rcpp::traits::input_parameter< const Eigen::MatrixXd& >::type x(xSEXP);
-    Rcpp::traits::input_parameter< const Eigen::MatrixXd& >::type c(cSEXP);
-    Rcpp::traits::input_parameter< const int >::type control(controlSEXP);
-    Rcpp::traits::input_parameter< const int >::type k(kSEXP);
-    Rcpp::traits::input_parameter< const double >::type level(levelSEXP);
-    Rcpp::traits::input_parameter< const bool >::type interval(intervalSEXP);
-    Rcpp::traits::input_parameter< const std::string >::type method(methodSEXP);
-    Rcpp::traits::input_parameter< const int >::type B(BSEXP);
-    Rcpp::traits::input_parameter< const int >::type nthreads(nthreadsSEXP);
-    Rcpp::traits::input_parameter< const double >::type th(thSEXP);
-    Rcpp::traits::input_parameter< const int >::type maxit(maxitSEXP);
-    Rcpp::traits::input_parameter< const double >::type abstol(abstolSEXP);
-    rcpp_result_gen = Rcpp::wrap(pairwise(x, c, control, k, level, interval, method, B, nthreads, th, maxit, abstol));
-    return rcpp_result_gen;
-END_RCPP
-}
 // test_GLM
[…]

### 2024-02-25 R/confint-methods.R
checkmate package adoption + pdf vignette (#4)
* checkmate package adoption wip

* checkmate adoption wip + bug fix in confint() argument setting for `ELMT`

* `validate_alpha()` removed

* argument validation using checkmate implemented paritally

* new vignette wip

* rnw test

* tex action test

* tex action wip

* tex action wip

* tex action wip

* tex action wip

* tex action wip

* tex action wip

* tex action wip

* tex action wip

* tex action wip

* tex action wip

* tex action wip

* compact vignette

* tex action wip

* sec 4.2 rnw done

* add .cls file to rbuildignore

* new vigne

@@ -39,7 +39,7 @@ setMethod("confint", "EL", function(object,
   }
   # Number of rows of the confidence interval matrix
   p <- length(idx)
-  level <- validate_level(level)
+  level <- assert_number(level, lower = 0, upper = 1, finite = TRUE)
   if (isTRUE(all.equal(level, 0))) {
     ci <- matrix(rep(est[idx], 2L), ncol = 2L)
     colnames(ci) <- c("lower", "upper")
@@ -95,13 +95,13 @@ setMethod("confint", "ELMT", function(object,
     stopifnot("Invalid `control` specified." = is(control, "ControlEL"))
   }
   method <- getMethodEL(object)
-  cv <- if (is.null(cv)) object@cv else validate_cv(cv, th)
   maxit <- control@maxit
   maxit_l <- control@maxit_l
   tol <- control@tol
   tol_l <- control@tol_l
   step <- control@step
   th <- control@th
+  cv <- if (is.null(cv)) object@cv else validate_cv(cv, th)
   nthreads <- control@nthreads
   w <- getWeights(object)
   estimates <- unlist(getEstimates(object))

### 2024-02-25 R/el_mean.R

@@ -55,26 +55,36 @@ el_mean <- function(x,
                     par,
                     weights = NULL,
                     control = el_control()) {
-  mm <- validate_x(x)
+  x <- as.matrix(x, rownames.force = TRUE)
+  assert_matrix(x,
+    mode = "numeric", any.missing = FALSE, all.missing = FALSE, min.rows = 2L,
+    min.cols = 1L
+  )
+  assert_numeric(x, finite = TRUE)
+  stopifnot(
+    "`x` must must have larger number of rows than columns." =
+      nrow(x) > ncol(x),
+    "`x` must have full column rank." = get_rank(x) == ncol(x)
+  )
   stopifnot(
     "`par` must be a finite numeric vector." =
       (isTRUE(is.numeric(par) && all(is.finite(par)))),
     "Invalid `control` specified." = (is(control, "ControlEL"))
   )
-  n <- nrow(mm)
-  p <- ncol(mm)
+  n <- nrow(x)
+  p <- ncol(x)
   if (length(par) != p) {
     stop(gettextf("Length of `par` must be %d.", p, domain = NA))
   }
   w <- validate_weights(weights, n)
   if (length(w) == 0L) {
-    est <- colMeans(mm)
+    est <- colMeans(x)
   } else {
-    est <- colSums(mm * w) / n
-    names(w) <- rownames(mm)
+    est <- colSums(x * w) / n
+    names(w) <- rownames(x)
   }
   out <- compute_EL(
-    "mean", par, mm, control@maxit_l, control@tol_l, control@th, w
+    "mean", par, x, control@maxit_l, control@tol_l, control@th, w
   )
   optim <- validate_optim(out$optim)
   names(optim$par) <- names(est)
@@ -86,10 +96,10 @@ el_mean <- function(x,
     )
   }
   new("EL",
-    optim = optim, logp = setNames(out$logp, rownames(mm)), logl = out$logl,
+    optim = optim, logp = setNames(out$logp, rownames(x)), logl = out$logl,
     loglr = out$loglr, statistic = out$statistic, df = p,
     pval = pchisq(out$statistic, df = p, lower.tail = FALSE), nobs = n,
     npar = p, weights = w, coefficients = est, method = "mean",
-    data = if (control@keep_data) mm else NULL, control = control
+    data = if (control@keep_data) x else NULL, control = control
   )
 }

### 2024-04-11 README.Rmd
pkgdown test
update unit tests

--- 2024-02-15 README.Rmd
@@
 
-# melt  
-<a href="https://docs.ropensci.org/melt/"><img src="man/figures/logo.png" align="right" height="139" /></a>
+# melt <a href="https://docs.ropensci.org/melt/"><img src="man/figures/logo.png" align="right" height="139" alt="melt website" /></a>
 
@@
 [![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
-[![R-CMD-check](https://github.com/ropensci/melt/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/ropensci/melt/actions/workflows/check-standard.yaml)
-[![Codecov test coverage](https://codecov.io/gh/ropensci/melt/branch/master/graph/badge.svg)](https://app.codecov.io/gh/ropensci/melt?branch=master)
+[![R-CMD-check](https://github.com/ropensci/melt/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ropensci/melt/actions/workflows/R-CMD-check.yaml)
+[![pkgcheck](https://github.com/ropensci/melt/workflows/pkgcheck/badge.svg)](https://github.com/ropensci/melt/actions?query=workflow%3Apkgcheck)
+[![Codecov test coverage](https://codecov.io/gh/ropensci/melt/branch/main/graph/badge.svg)](https://app.codecov.io/gh/ropensci/melt?branch=main)
 [![CRAN status](https://www.r-pkg.org/badges/version/melt)](https://CRAN.R-project.org/package=melt)
@@
 * `confreg()` computes confidence region for model parameters.
-* `elt()` tests a linear hypothesis.
+* `elt()` tests a hypothesis with various calibration options.
 * `elmt()` performs multiple testing simultaneously.

### 2024-04-11 R/elt-methods.R

@@ -45,7 +45,7 @@ setMethod("elt", "EL", function(object,
       pg <- -an * colMeans(g)
       g <- rbind(g, pg)
       out <- compute_generic_EL(g, maxit_l, tol_l, th, w)
-      out$optim <- c(par = par, out$optim)
+      out$optim$par <- par
     }
     optim <- validate_optim(out$optim)
     names(optim$par) <- pnames
@@ -139,7 +139,7 @@ setMethod("elt", "QGLM", function(object,
       pg <- -an * colMeans(g)
       g <- rbind(g, pg)
       out <- compute_generic_EL(g, maxit_l, tol_l, th, w)
-      out$optim <- c(par = par, out$optim)
+      out$optim$par <- par
     }
     optim <- validate_optim(out$optim)
     names(optim$par) <- nm
@@ -226,7 +226,7 @@ setMethod("elt", "SD", function(object,
       pg <- -an * colMeans(g)
       g <- rbind(g, pg)
       out <- compute_generic_EL(g, maxit_l, tol_l, th, w)
-      out$optim <- c(par = par, out$optim)
+      out$optim$par <- par
     }
     optim <- validate_optim(out$optim)
     names(optim$par) <- pnames

### 2024-05-16 src/compute_bootstrap_calibration.cpp
bump to v1.11.4

@@ -41,7 +41,7 @@ Rcpp::NumericVector compute_bootstrap_calibration(
       x0_fn = transform_x_fn(method);
   const Eigen::MatrixXd g0 = g_fn(x0_fn(x, par, est), par);
   // initialize seed
-  dqrng::xoshiro256plus gen(seed);
+  dqrng::xoshiro256plusplus gen(seed);
   // discrete uniform distribution
   boost::random::uniform_int_distribution<> u(0, g0.rows() - 1);
   // bootstrap statistics
@@ -52,7 +52,7 @@ Rcpp::NumericVector compute_bootstrap_calibration(
   {
   #endif
     // make thread local copy of gen
-    dqrng::xoshiro256plus lgen(gen);
+    dqrng::xoshiro256plusplus lgen(gen);
     #ifdef _OPENMP
     // advance gen by 1 ... nthreads jumps
     lgen.jump(omp_get_thread_num() + 1);
